### 2016-07-06 Useful (anit)Commutative relationships.tex

\documentclass{article}

\usepackage{float}  % For H in figures
\usepackage{amsmath} % For math
\usepackage{amssymb}
\usepackage{mathtools} % dcases*, see https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics#The_cases_environment
\numberwithin{equation}{subsection} % have the enumeration go to the subsection level.
									% See:https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics
\usepackage{graphicx}   % need for figures
\usepackage{cite} % For bibligraphy
\usepackage{fancyref} % For lazy reference \fref
\usepackage{hyperref} % For hyperlink everything.
\usepackage{CJKutf8} % For Chinese characters

\title{Useful (anti)Commutation Relation}
\date{\today}
\author{\begin{CJK}{UTF8}{gbsn}
陈鸿翔
\end{CJK}(Taper)}

\begin{document}

\maketitle

\abstract{
none
}

\tableofcontents

\section{General}
\begin{align}
[A,B]:=AB-BA \\
\{A,B\}:=AB+BA\\
[AB,C] &= A [B,C] + [A,C]B\\
		&= A\{B,C\} - \{A,C\} B\\
[C,AB] &= -[AB,C]\\
		&=A [C,B] + [C,A]B \\
       &=\{C,A\} B - A\{C,B\}
\end{align}

\section{2nd Quantization}

\subsection{Boson}

\begin{align}
n_{\alpha} := a_{\alpha}^{\dagger}a_{\alpha}\\
[a_{\alpha},a_{\beta}]=[a^{\dagger}_{\alpha},a^{\dagger}_{\beta}]=0\\
[a_{\alpha},a^{\dagger}_{\beta}]=\delta_{\alpha\beta}
\end{align}

\subsection{Fermion}
\begin{align}
n_{\alpha} := c_{\alpha}^{\dagger}c_{\alpha}\\
\{c_{\alpha},c_{\beta}\} = \{c_{\alpha}^{\dagger},c_{\beta}^{\dagger}\} = 0\\
\{c_{\alpha},c^{\dagger}_{\beta}\}=\delta_{\alpha\beta} \\
\{n_{\alpha},c_{\alpha}\} = c_{\alpha} \\
\{n_{\alpha},c^{\dagger}_{\alpha}\} = c^{\dagger}_{\alpha}
\end{align}

\begin{thebibliography}{1}
	\bibitem
	{wiki:Matrix} Wikipedia contributors, "Matrix mechanics," 
	Wikipedia, The Free Encyclopedia, 
	\url{https://en.wikipedia.org/wiki/Matrix_mechanics}
	(accessed July 2, 2015).
	\bibitem
	{Green} Green, H.S. 1965. Matrix Mechanics. Groningen, 
	The Netherlands: P. Noordhoff Ltd.
	\bibitem{Wrong}
	H. Fritzsch. You Are Wrong, Mr Einstein! World Scientific Publishing Company.
	\bibitem{Griff}
	D.J Griffiths. Introduction to Quantum Mechanics, 3rd edition.
	Pearson Prentice Hall.
	\bibitem{S-L_theory}
	Sturm-Liouville theory. Encyclopedia of Mathematics. 
	\url{http://www.encyclopediaofmath.org/index.php?title=Sturm-Liouville_theory&oldid=23061}
	(Accessed July 14, 2015)
\end{thebibliography}
\end{document}