### 2016-09-06 G.R/G.R.tex
Initiating new fields: general relativity and quantum gravity.

% The entire content of this work (including the source code
% for TeX files and the generated PDF documents) by 
% Hongxiang Chen (nicknamed we.taper, or just Taper) is
% licensed under a 
% Creative Commons Attribution-NonCommercial-ShareAlike 4.0 
% International License (Link to the complete license text:
% http://creativecommons.org/licenses/by-nc-sa/4.0/).
\documentclass{article}

\usepackage{float}  % For H in figures
\usepackage{amsmath} % For math
\usepackage{amssymb}
\usepackage{mathrsfs}
% Followings are for the special character: differential "d".
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}
\numberwithin{equation}{subsection} % have the enumeration go to the subsection level.
                                    % See:https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics
\usepackage{graphicx}   % need for figures
\usepackage{cite} % need for bibligraphy.
\usepackage[unicode]{hyperref}  % make every cite a link
\usepackage{CJKutf8} % For Chinese characters
\usepackage{fancyref} % For easy adding figure,equation etc in reference. Use \fref or \Fref instead of \ref
\usepackage{braket} %http://tex.stackexchange.com/questions/214728/braket-notation-in-latex

% Following is for theorems etc environments
% http://tex.stackexchange.com/questions/45817/theorem-definition-lemma-problem-numbering && https://en.wikibooks.org/wiki/LaTeX/Theorems
\usepackage{amsthm}
\newtheorem{defi}{Definition}[section]
\newtheorem{thm}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{prop}{Proposition}[section]
\newtheorem{coro}{Corollary}[section]
\theoremstyle{definition}
\newtheorem{ex}{Example}[section]

% A list of nomenclatures.
\usepackage{nomencl}
\makenomenclature

\title{Notes for General Relativity}
\date{\today}
\author{Taper}

\begin{document}

\maketitle
\abstract{
    (None)
}
\tableofcontents
\section{The Principle of Relativity}
\label{sec:Principle_of_Relativity}
Landau's book \cite{landau} describes the speed of light, as a speed of the
maximum velocity of propagation of interaction. He perceives velocity as
describing the propagation of interaction.

In another aspect, I think that time does not exist. I perceive time as
a agent of the environmental effect, communicating information and 
coordinating movement between the system under consideration and its 
envrionment. Therefore, a single existance is eternal by nature. Perhaps 
before the universe originates, time is a meaningless concept. In my 
view, the light is the only reliable tool to serve function of
communication and coordination.

Fram either perspective, the speed of light is inherently constent.
However, from my view, it is not clear why this speed should be a maximum.

\subsection{Invariance of Interval}
\label{sec:Invariance of Interval}

The Landau's book\cite{landau}, introduces a

\begin{thebibliography}{1}
    \bibitem{landau} The Classical Theory of Fields
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}