### 2016-10-05 Notes about TR symmetry/Notes about TR symmetry.tex
New drafts: something calculation about spin operators. Plans on new notes about TR symmetry for condensed matter physics.

% The entire content of this work (including the source code
% for TeX files and the generated PDF documents) by 
% Hongxiang Chen (nicknamed we.taper, or just Taper) is
% licensed under a 
% Creative Commons Attribution-NonCommercial-ShareAlike 4.0 
% International License (Link to the complete license text:
% http://creativecommons.org/licenses/by-nc-sa/4.0/).
\documentclass{article}

\usepackage{float}  % For H in figures
\usepackage{amsmath} % For math
\usepackage{amssymb}
\usepackage{bbm} % for numbers within mathbb
\usepackage{mathrsfs} % For \mathscr{ABC}
% Followings are for the special character: differential "d".
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}
\numberwithin{equation}{subsection} % have the enumeration go to the subsection level.
                                    % See:https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics
\usepackage{graphicx}   % need for figures
\usepackage{cite} % need for bibligraphy.
\usepackage[unicode]{hyperref}  % make every cite a link
\usepackage{CJKutf8} % For Chinese characters
\usepackage{fancyref} % For easy adding figure,equation etc in reference. Use \fref or \Fref instead of \ref
\usepackage{braket} %http://tex.stackexchange.com/questions/214728/braket-notation-in-latex

% Following is for theorems etc environments
% http://tex.stackexchange.com/questions/45817/theorem-definition-lemma-problem-numbering && https://en.wikibooks.org/wiki/LaTeX/Theorems
\usepackage{amsthm}
\newtheorem{defi}{Definition}[section]
\newtheorem{thm}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{prop}{Proposition}[section]
\newtheorem{coro}{Corollary}[section]
\theoremstyle{definition}
\newtheorem{ex}{Example}[section]

% A list of nomenclatures.
\usepackage{nomencl}
\makenomenclature

% For drawing diagrams with arrows
\usepackage[all]{xy}

\title{Notes about TR Symmetry}
\date{\today}
\author{Taper}

\begin{document}

\maketitle
\abstract{
This is a note about TR symmetry, with its application in the theory of
condensed matter physics. The two reference book are the classic by
Sakurai \cite{sakurai}, and the other by Bernevig \cite{bernevig}.
}
\tableofcontents

\section{Review of Spin theory}
\label{sec:Review-of-Spin-theory}
Outline:
\begin{enumerate}
    \item The representation of infinitesimal operator.
    \item From traditional rotational operators to commutation
        relationship
    \item From commutators to their eigenvalues.
    \item Summarize some further results regarding the matrix
        representation of those operators.
\end{enumerate}

\section{Review of Symmetry Operator}
\label{sec:Review-of-Symmetry-Operator}
\begin{enumerate}
    \item The invariance of probability and Wigner's theorem
    \item The values of measurement before and after transformation
\end{enumerate}

\section{Derive some properties of TR operator}

\section{Kramer's theorem and the lifting of degenracy}

\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    \bibitem{sakurai} Modern Quantum Mechanics. J.J. Sakurai.
    \bibitem{bernevig} Introduction to Topological Insulators. Bernevig.
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}