### 2016-10-26 Draft/sections/misc-math.tex

\subsection{\texorpdfstring{$\int_0^\infty \frac{\sin(x)}{x}$}{}}
From \href{http://math.stackexchange.com/questions/5248/solving-the-integral-int-0-infty-frac-sinxx-dx-frac-pi2}{Math.SE}. 
By \href{http://math.stackexchange.com/users/1102/aryabhata}{Aryabhata}.

I believe this can also be solved using double integrals.

It is possible (if I remember correctly) to justify switching the order of integration to give the equality:

$$\int_{0}^{\infty} \Bigg(\int_{0}^{\infty} e^{-xy} \sin x \,dy \Bigg)\, dx = \int_{0}^{\infty} \Bigg(\int_{0}^{\infty} e^{-xy} \sin x \,dx \Bigg)\,dy$$
Notice that
$$\int_{0}^{\infty} e^{-xy} \sin x\,dy = \frac{\sin x}{x}$$

This leads us to

$$\int_{0}^{\infty} \Big(\frac{\sin x}{x} \Big) \,dx = \int_{0}^{\infty} \Bigg(\int_{0}^{\infty} e^{-xy} \sin x \,dx \Bigg)\,dy$$
Now the right hand side can be found easily, using integration by parts.

\begin{align*}
I &= \int e^{-xy} \sin x \,dx = -e^{-xy}{\cos x} - y \int e^{-xy} \cos x \, dx\\
&= -e^{-xy}{\cos x} - y \Big(e^{-xy}\sin x + y \int e^{-xy} \sin x \,dx \Big)\\
&= \frac{-ye^{-xy}\sin x - e^{-xy}\cos x}{1+y^2}.
\end{align*}
Thus $$\int_{0}^{\infty} e^{-xy} \sin x \,dx = \frac{1}{1+y^2}$$
Thus $$\int_{0}^{\infty} \Big(\frac{\sin x}{x} \Big) \,dx = \int_{0}^{\infty}\frac{1}{1+y^2}\,dy = \frac{\pi}{2}.$$

\subsection{Bayes' theorem}
\label{sec:Bayes-theorem}
I am quite unfamiliar with this formula, so I decided to make a note
here for future reference.

This formula starts from the definition of conditional probability:
\begin{align}
P(A|B)\equiv \frac{P(A\&B)}{P(B)}
\end{align}
this definition if quite intuitively pleasing if $P(B)$ is multiplied
to the left side. Then one easily deduce that:
\begin{thm}[Bayes' theorem] \begin{align}
   	P(A|B)=\frac{P(B|A)P(A)}{P(B)}
   	\end{align} \end{thm}

We have also an extended form. Suppose the event space is partitioned
into $\{A_i\}$, then we have (also easily proved):
\begin{align}
P(A_i|B)=\frac{P(B|A_i)P(A_i)}{\sum_j P(B|A_j)P(A_j)}
\end{align}
Here's a good reading \href{http://plato.stanford.edu/entries/bayes-theorem/}{Bayes' Theorem in SEP}.

\subsection{System of Differential Equations}
This is a smsall note of \cite{DETA}.

pp. 266.

\begin{defi}
   	$\mathbf{x(t)}$ is a vector whose elements are $x_i(t)$.
   	$ \frac{d}{d t}$ acts on vector $\mathbf{x}$ element-wise.
   	$\dot{\mathbf{x}}$ is abbrevation for $\frac{d}{d t}\mathbf{x}$
\end{defi}

pp. 291.

\begin{thm}[Existence-uniqueness theorem]
   	There exists one, and only one, solution of the initial-value
   	problem

   	\begin{align}
   	\dot{\mathbf{x}}=\mathbf{A}\mathbf{x}\text{, }&
   	\mathbf{x}(t_0) = \mathbf{x}^0 = 
   	\left(
   	\begin{array}{c}
   	x^0_1\\
   	x^0_2\\
   	\cdots
   	\end{array} 
   	\right)
   	\end{align}

   	Moreover, this solution exists for $-\infty\langle t\langle \infty$.
\end{thm}
\begin{remark}
   	By this, any non-trivial solution $\mathbf{x}(t)\neq 0$ at any
   	time $t$. Also notice that the elements of $\mathbf{A}$ are just
   	numbers.
\end{remark}

\begin{thm}
   	The dimension of the space $\mathbf{V}$ of all solutions of the
   	homogeneous linear system of differential equations:
   	\begin{align}
   	\frac{d\mathbf{x}}{dt}=\mathbf{Ax}
   	\end{align}
   	is $n$, i.e. the dimension of vector $\mathbf{x}$.
\end{thm} 
\subsection{ODE by Arnold}
sec. 14
\begin{defi}
   	\begin{align}
   	\label{eq:e^A}
   	e^A &= I + A + \frac{A^2}{2!} + \frac{A^3}{3!}\\
   	\text{or}& \nonumber \\
   	e^A &= \lim_{n\to \infty}(I+\frac{A}{n})^n
   	\end{align}
   	where $I$ is the identity matrix.
\end{defi}
Equivalance of the two definition will be addressed in the Theorem on
pp. 165.

Important theorems:
\begin{thm}[pp. 158]
   	The series $e^A$ converges for any $A$ uniformly on each set
   	$X=\{A:||A||\leq a\}$, $a\in \mathbb{R}$.
\end{thm}
\begin{thm}[pp. 160]
   	$$e^{At} = H^t$$
   	where $H^t$ is the translation operator which sends every polynomial
   	$p(x)$ into $p(x+t)$.
\end{thm}
\begin{thm}[pp. 163]
   	$$\frac{d}{dt} e^{tA} = Ae^{tA}$$
\end{thm}
\begin{thm}[Fundamental Theorem of the Theory of Linear Equations with
   	Constant Coefficients]
   	The solution of:
   	\begin{align}
   	\label{eq:fund_thm_of_linear_eqs_const_coef}
   	\dot{\mathbf{x}} = A\mathbf{x}
   	\end{align}
   	with initial condition $\phi(0) = \mathbf{x}_0$ is
   	\begin{align}
   	\mathbf{\phi}(t) = e^{tA}\mathbf{x}_0
   	\end{align}
\end{thm}

Practically solution to
$$ \dot{\mathbf{x}} = A\mathbf{x}$$
(pp. 173, Sec 17)
(Assuming $A$ is diagonalizable.)
\begin{itemize}
   	\item Find the eigenvectors $\xi_1,\cdots ,\xi_n$ and eigenvalues
   	$\lambda_1,\cdots ,\lambda_n$. Use them as basis.
   	\item Expand the initial condition in the new basis.
   	\begin{align}
   	\mathbf{x}_0=\sum_{k=1}^{n} C_k\xi_k
   	\end{align}
   	\item Then $\phi(t) = \sum_{k=1}^n C_k e^{\lambda_k t}\xi_k$
\end{itemize}

\subsection{Why 0/0 is undefined?}
If we suppose
$$ \frac{0}{0}= \triangle $$
Consider the following derivation:
\begin{align}
\frac{0}{0} \cdot 1 &= \triangle \cdot 1 = \triangle \\
0 \cdot \frac{1}{0} &= \triangle\\
\Rightarrow \triangle &= 0
\end{align}
This is already bad enough. And we are forced to define $\frac{1}{0}$.
Let $\frac{1}{0} = \square$, which literally means $1=0\cdot \square = 0$.
This is disastrous.

Alternatively, we could let
\begin{enumerate}
   	\item Let $\frac{1}{0}$ be undefined.
   	\item Or let $\frac{1}{0} =
   	\infty$.
   	\item Or, let $\frac{a}{b}\cdot c= a\cdot \frac{c}{b}$ be not 
   	true when $b=0$.
\end{enumerate}
The third idea is disastrous for algebraic manipulation. \footnote{
   	Or more speicifically, it is a disaster for field theory.
}
The first idea is not good. Since defining $\frac{1}{0}=\infty$ turns
out to be very useful in both mathematics and physics. Actully, in
physics it is common practice to set $\frac{a}{0}=\pm\infty$ for
any nonzero number $a$, where the sign of $\infty$ is determined by 
the sign of $a$.
The second idea is okey. But then we are faced with a serious problem.
We have to define $\triangle \equiv 0 \cdot \infty$

$\triangle \cdot 2 = \triangle$, What will be of $\triangle + 1$?