### 2016-10-26 Draft/sections/misc-other.tex

\subsection{About Matlab}
\label{sec:About-Matlab}
\paragraph{On using $A/B$ or $A\\B$}
Matlab's matrix divide function \texttt{mrdivide}, \texttt{mldivide}
, abbrevated $A/B$ and $A\\B$, are different and should be used with
caution. One can check that:
\begin{center}
    $A\cdot D^{-1}$ is equivalent to $A/D$, $D^{-1}\cdot A$ is
    equivalent to $D\backslash A$.
\end{center}
Also, it is straightforward to prove that the two are equal if and only if $A$ commutes with $D$.

\subsection{The Mathematical Theory of Communication}
\label{sec:The_Mathematical_Theory_of_Communication}
\subsubsection{Introduction}

\paragraph{What is information}
In this part, it is implied that \textit{information} in this work
does not carry the usual sense in people's daily life. The semantic
aspect of a message is considered to be irrelavant, for purpose
of generality of the design of communication systems.

\paragraph{Measure of information} 
% TODO Read the paper of Hartley to understand what is information:
% Hartley, R. V. L., "Transmission of Information," Bell System Technical
% Journal, July 1928, p. 535

Then it refers to a paper by Hartley to substantiate the use of
\begin{align}
\label{eq:measure_of_information}
S = \log(M)
\end{align}
as a measure of information. More specifically, we assume we have a
set of possible messages. Then $M$ is the cardinality of this set.
Then $S$ is a measure of the information produced when one message
is chosen from the set. Once again, we regard all choices 
being equally possible.

Note that the base of logrithm in \ref{eq:measure_of_information}
is undefined. Choosing a base constituting choosing a unit of
the measure. Two such measures, when calculated in different units,
are related by a simple constant.

Conventionally, a base $2$ is chosen. The resulted unit is called
bits. If the base $10$ is chosen, then the units may be called
decimal digits. If the base $e$ is chosen, then the units is
called natural units.

Also, the author lists several points to illustrate the
convenience of this measure.

\paragraph{Communication systems}
Next the author defines the necessary components of a \textit{
   	communication system}, and categorizes it into discrete systems,
continuous systems and mixed systems.

\subsubsection{Discrete Noiseless Systems}

\paragraph{Discrete Noiseless Channel}
\label{sec:Discrete Noiseless Channel}
This part deals with another measure, the measure of the capacity
of a channel to transmit information. It defines the capacity of a
discrete channel as:
\begin{align}
\label{eq:capacity_of_disc_chan}
C\equiv \lim_{T\to \infty} \frac{\log N(T)}{T}
\end{align}
where $N(T)$ is the number of allowed signals of duration $T$.
Several examples are given with formula for $C$ in each particular
example.

\paragraph{Discrete Source of Information}
Next, it proceeds to discuss the statistical property of the source
of information. Pointing out that a statistical knowledge of the
source of information can help peopole craft special protocals to
reduce the required capacity of the channel, the article gradually
focuses on the statistical property of sourcs. It profess that while a
discrete source could be represented by a statistical source, a
statistical process can also be considered a discrete source. The
second claim is substatiated by several exmaples.

In one example of natural language, the article defines a 
\textit{n-gram} case to produce natural language from statistical
information.

\paragraph{Series of Approximations to English}
As the title suggests, this part illustrate two serial levels of
steps to approximate the English language using statistical knowledge
of appearance of alphabets (the first method) and words (the second
example). The article claims that "a sufficiently complex stochastic 
process will give a satisfactory representation of a discrete source".
Although I am largely against this juvenile view.

\paragraph{Graphical Representation of a Markoff Process}
Then the article mentions a graphical way to represent the aforementioned
approximation process, and gives three examples on page 46.

\paragraph{Ergodic and Mixed Sources}
Now the article comes to a special type of stochastic process, ergodic
processes. A rough idea of "ergodic" is given in page 45. The idea is
so important that I felt compeled to present it here:

"In an ergodic process every sequence produced by
the process is the same in statistical properties. Thus the letter
frequencies, digram frequencies, etc., obtained from particular
sequences, will, as the lengths of the sequences increase, approach
definite limits independent of the particular sequence.
Actually
this is not true of every sequence but the set for which it is false
has probability zero. Roughly the ergodic property means 
\textbf{statistical homogeneity}."

Next, the article claims that artificial languages given in previous
examples are ergodic, because the corresponding graph does not
have two properties: they does not comprise two or more \textit{isolated
   	parts}, and they $gcd$ of the lengths of all \textit{circuits }is one.
The precise meaning is listed in page 47. Roughly, an analogy made by
myself helps to
catch the points. If we picture an stochastic process as a connected 
area, then isolated parts are its connected components, whereas the
circuit are the recurrent pattterns.

Naturally , a stochastic process may exhibit a mixed behavir, in which
there are several different sources $L_1,L_2,L_3,\cdots$, which are
each of homogeneous, i.e. ergodic, statistical structure. This is
discussed following the introduction of ergodicity in page 48.

Then the article declare that except in special cases, ergodicity is
always assumed. This purpose is analogous to that of in statistical
physics, to "identify averages along a sequence with averages over 
the ensemble of possible sequences", with "the probability of a
discrepancy being zero".

Lastly, the article mentions a fact regarding the equilibrium of
the system. A process is called stationary, if it satisfies a
equilibrium condition:
\begin{align}
\label{eq:Ergodic_and_Mixed_States:equil_condition}
P_j = \sum_i P_i \cdot P_i(j)
\end{align}
where $P_j$ is the probability of being in state $j$, and $P_i(j)$ is
the transition probability from $i$ to $j$.
The fact is that ergodic process is, in a sense, always stationary.

\paragraph{The Entropy of an Information Source}
This part first defines the entropy of a discrete source of finite state
to be:
\begin{align}
\label{eq:entropy_of_discrete_infor_source}
H \equiv& \sum_i P_i H_i\\
=& -\sum_{i,j} P_i p_i(j) \log(p_i(j)) \nonumber
\end{align}
It is "the entropy of the source per symbol of text". Another
definition for entropy per second is also listed.

Following this definition are some theorems, which I consider to be
the most essential and influential part of the whole book (although
I have not yet read the whole book). They are:

\begin{thm}[Theorem 3 on page 55]
   	Given any $\epsilon >0$ and $\delta >0$, we can find an $N_0$
   	such that the sequences of any legnth $N\geq N_0$ fall into
   	two class:
   	\begin{itemize}
   		\item A set whose total probability is less than $\epsilon$.
   		\item The remainder, all of whose members have probabilities
   		satisfying the inequality
   		\begin{align}
   		\label{eq:Entropy_of_info_source:thm3}
   		\left| \frac{\log p^{-1}}{N}-H\right| < \delta
   		\end{align}
   	\end{itemize}
\end{thm}
"In other words we are almost certain to have 
$\frac{\log p^{-1}}{N}$ very close to $H$ when $N$ is large."
For me, this reads quite like the \textit{second law of
   	thermaldynamics}.

% TODO having trouble understanding the Theorem 4 on page 55.