### 2016-11-02 Notes of various papers/Kitaev. Periodic table for topological insulators and superconductors/notes.tex

% The entire content of this work (including the source code
% for TeX files and the generated PDF documents) by 
% Hongxiang Chen (nicknamed we.taper, or just Taper) is
% licensed under a 
% Creative Commons Attribution-NonCommercial-ShareAlike 4.0 
% International License (Link to the complete license text:
% http://creativecommons.org/licenses/by-nc-sa/4.0/).
\documentclass{article}

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physicist-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}

\makenomenclature

\title{Notes about Periodic table for topological insulators and superconductors}
\date{\today}
\author{Taper}

\begin{document}

\maketitle
\abstract{
As title suggests.
}
\tableofcontents
From the start: for \textbf{Gapped free-fermion} system, he mentions
the classification hitherto available. See
Table~\ref{tab:classification}.
\begin{table}[ht]
    \centering
    \caption{Classification}
    \label{tab:classification}
    \vspace{+0.5pt}
    \rowcolors{1}{white}{lightgrey}
    \tabulinesep=1.2mm
    \begin{tabu}{  X[l] X[l] X[l] X[l] }
        \textit{Symmetry} & \textit{Spatial Dimension} &
        \textit{Result} & \textit{Other Keywords} \\
        \hline
        T & 0 & An intger: the number of particle-occupied Kramers
        doublet states & \\
        T & 1 & None & \\
        T & 2 & $\mathbb{Z}_2$ & \\
        T & 3 & $\mathbb{Z}_2$ & $3D$ crystals have additional
        $3\mathbb{Z}_2$ invariant$ \Rightarrow$ "weak topological
        insulators \\
        Q(?) & 2 & Characterized by $\mu$ in units of $e^2/h$ & TKNN \\
        Q(?) & even  d & Topological invariant ($k$-th Chern number) \\
        Q(?) & 0 &  number of single-particle states with negative
energy ($E< E_F = 0$), which are filled with electrons. \\
        T\& Q \\
        No T \& No Q & 0 & $\mathbb{Z}_2$ \\
        No T \& No Q & 1 &$\mathbb{Z}_2$ &"majorana chain"  \\
        No T \& No Q & 2 & Topological number is integer. & Even-odd
        effects. \\
    \bottomrule
    \end{tabu}
\end{table}

\paragraph{Robustness in classification}
\begin{quote}
    ..., the topological numbers are insensitive to disorder and can even
    be defined without the spectral gap assumption, provided the
    eigenstates are localized.
\end{quote}

\paragraph{Equivalence of Hamiltonian} 
\begin{quote}
    ... Two Hamiltonians belong to the same phase if they can be
    continuously transformed one to the other while maintaining
    the energy gap or localization; we will elaborate on that
    later. 
\end{quote}

\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    \bibitem{kitaev-Periodic} Alexei Kitaev. Periodic table for
    topological insulators and superconductors.

\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}