### 2016-11-13 Lecture note - Group Theory in Physics/sections/20161107.tex

\subsection{Finding Representation}
\label{sec:Finding-Representation}
One important work in group theory is to find all its irreducible
representations. The book \cite{book} suggests we first find the
characters of irreducible representations, and list them in a
character table, and then find the matrix of those irreducible
representations.

He also notes that, a unfaithful representation of a group is
equivalent to a faithful representation of a certain quotient group.
Since quotient group is less complex than the original group, finding
irreducible representation of normal subgroups is very useful for
analysing a group.

\begin{fact}
    Also, a the direct product of two irreducible representations is also
    a representation. If the dimension of one of the representation is
    $1$, then their product is still a irreducible representation.
\end{fact}
\subsection{Self-conjugate representation and Real Representation}
\label{sec:Self-conjugate-representation-and-Real-Representation}

It's easy to see that for any representation $D$, conjugating ever
matrix $D(g)^*$ will give another representation. We define
\begin{defi}[Conjugate Representation]
\nomenclature{Conjugate Representation}{\nomrefpage.}
    The conjugate representation of a given representation $D$ is obtained
    by complex conjugate all its matrices $D(g)^*,\,g\in G$.
\end{defi}
\begin{defi}[Self-conjugate Representation]
\nomenclature{Self-conjugate Representation}{\nomrefpage.}
    A representation $D$ is called self-conjugate if and only if it is
    equivalent (but not necessarily equal) to its conjugate
    representation $D^*$.
\end{defi}
\begin{fact}
    The character of a self-conjugate representation is a real
    function. Since (assume $XD(g)X^{-1}=D^*(g)$):
    $$
        \Tr(XD(g)X^{-1}) = \Tr(X^{-1}XD(g)) = \Tr(D(g)) = \Tr(D^*(g))
        = \Tr(D(g))^*
    $$
\end{fact}

We have a similar concept of real representation:
\begin{defi}[Real representation]
\nomenclature{Real representation}{\nomrefpage.}
    A representation $D$ is called real if and only if it is
    equivalent to a representation $D^\text{real}$, such that for any
    $g\in G$, the matrix $D^\text{real}(g)$ is real.
\end{defi}

Observe that a real representation is obviously self-conjugate, but a
self-conjugate representation is not necessarily real. We have two
indicator for its exact classifications.

Firstly,
\begin{thm}
    Let $D$ be a unitary irreducible representation of a finite group.
    Let $D^*$ be its conjugate representation. Assume that $D$ is
    self-conjugate. Let $X$ be the matrix that relates the two. Then
    $X$ is either symmetric or anti-symmetric. $X$ is symmetric if and
    only if $D$ is real. $X$ is anti-symmetric if and only if $D$ is
    self-conjugate but is not real.
\end{thm}

Secondly, 
\begin{thm}
    for any irreducible representation $D$ of finite group:
    \begin{align}
        \frac{1}{|G|}\sum_{g\in G} \chi(g^2) = \begin{cases}
            1, & \text{$D$ is real} \\
            -1, & \text{$D$ is self-conjugate but is not real} \\
            0, & \text{$D$ is not self-conjugate}
        \end{cases}
    \end{align}
\end{thm}

The proof of these two facts are presented in page 68 of \cite{book}.
Note that there is more general version of the second fact, called
\textit{Frobenius–Schur indicator}, see
\href{https://en.wikipedia.org/wiki/Frobenius%E2%80%93Schur_indicator}
{Wikipedia page}.

\subsection{Subduced Representation}
\label{sec:Subduced-Representation}
\marginpar{To be finished.}
Let $G$ be a group. Let $\mathcal{C}_a$ denotes it conjugacy classes.
Let $D^j$ be its irreducible representations, with dimension $m_j$,
character $\chi^j_a$. Assume it has a subgroup $H=\{T_1=E,
T_2,\cdots,T_h\}$, where $h=|H|$. Let $n=|G|/h$ be its index. Denotes
its coset by $R_rH$. With $R_1:=E$, any element in $G$ can be denotes
as $R_r T_t$ (assume $R_r$ is chosen). Let $\bar{\mathcal{C}}_\beta$
denotes conjugacy classes of the subgroup $H$, with character
$\bar{\chi}^k_\beta$, dimension $\bar{m}_k$.

% \begin{thm}[Frobenius Theorem]
%     % TODO
% \end{thm}
% \begin{proof}
%     Assume that $D^j(T_t)$ can be decomposed into:
%     \begin{align}
%         X^{-1}D^j(T_t)X &= \oplus_k a_{jk}\bar{D}^k(T_t) \\
%         m_j = \sum_k a_{jk} \bar{m}_k
%     \end{align}
%     It is easy to prove that
%     \begin{equation}
%         a_{jk} = \frac{1}{|H|} \sum_{T_t\in H} \bar{\chi}^k(T_t)^*
%         \chi^j(Tt) 
%         = \frac{1}{|H|} \sum_\beta \bar{n}_\beta
%         (\bar{\chi}^k_\beta)^* \chi^j_\beta
%     \end{equation}
%     (Notice that, since character is invariant in conjugate class, we
%     can also label the conjugacy class of $\chi^j_\alpha$ with
%     $\beta$, where $\beta$ labels the conjugacy of subgroup.)
% \end{proof}