### 2016-11-13 Lecture note - Group Theory in Physics/sections/point-group.tex

Here I collect the result related to point group in physical space.
\begin{defi}[Point group]
\nomenclature{Point group}{\nomrefpage.}
    A group is called a point group if it contains of rotations
    that leaves a common point unchanged.
\end{defi}
\begin{defi}[Proper point group]
\nomenclature{Proper point group}{\nomrefpage.}
    A point group is called proper if its elements keep the orientation of the
    volumn element.
\end{defi}
\begin{defi}[Improper point group]
\nomenclature{Improper point group}{\nomrefpage.}
    Improper point group is a point group that contains both improper
    and proper rotations.
\end{defi}
\begin{defi}[Spatial inversion $\sigma$]
\nomenclature{Spatial inversion $\sigma$}{\nomrefpage.}

\end{defi}
\begin{remark}
    Let $G$ be a improper point group.
    An improper rotation in it can been seen as the direct product of a
    proper rotation and a spatial inversion $\sigma$. (Note that
    $\sigma$ commutes with any rotation, and $\sigma^2=\id$.)
    Therefore, a improper point group will always contain some proper
    rotation by $\sigma^2=\id$, and of course, the $e$ is always
    proper. Also, the subset of proper rotation will form a subgroup
    $H$. And such division will make $G/H \cong \mathbb{Z}_2$.
\end{remark}
However, a improper point group does not necessarily contains
$\sigma$. For example, assuming a proper point group has a subgroup
$H$ of index $2$, then multiply every elements of the coset by
$\sigma$ will gives us a group denoted $\sigma H$. Then $H\otimes
\sigma H$ gives us a improper point group that does not contain
$\sigma$.

\begin{defi}[P-type and I-type improper point group]
\nomenclature{P-type and I-type improper point group}{\nomrefpage.}
    A P-type improper point group is the improper point group that
    does not contain $\sigma$, spatial inversion. On the other hand,
    if it does contain $\sigma$, it is called an I-type improper point
    group.
\end{defi}

\subsection{Rotation \texorpdfstring{$C_n$}{}}
\label{sec:C_n}
$C_n$ is isomorphic to $\mathbb{Z}_n$, here the $C$ emphasize that it
represent the group generated by the rotation of $2\pi/n$. This group
is abelian.

\paragraph{Conjugacy Classes} Obvious each element form a conjugacy
class on its only. Moreover, when $n$ is even, the conjugacy class of
rotation $\pi$ is self-conjugate.

\paragraph{Invariant Subgroup} When $n$ is prime, there is no
nontrivial invariant subgroup. When $n$ can be factorized as $n_1n_2$,
both not equals $1$, then $C_n$ is the product of $C_m$ and $C_n$,
both are invariant subgroup.

\paragraph{Irreducible representation} Since it has $n$ elements and
$n$ conjugacy classes, with $\sum_{j=1}^n d_j^2 = n$, each irreducible
representation is $1$-dimensional. Let $R$ denotes the generator of
rotation by $2\pi/n$, then by calculating $D^j(R)^n=D^j(1)$, one can
easily find that for each irreducible representation $j$:
\begin{equation}
    D^j(R) = \exp(-i2\pi j/N)
\end{equation}
where $j=0,1,\cdots, n$. Since $R$ is its generator, the rest of the
representation is determined. Hence we have all the irreducible
representations of $C_n$. The table of characters may be found on page
65 of \cite{book}.

Note that since all representations are $1$-dimensional, their
products will keep being irreducible. So having two set of irreducible
representations of $C_n$ and $C_m$, will give us all the irreducible
representations of $C_{nm}$.
\subsection{Dihedral Group}
\label{sec:Dihedral-Group}
\begin{defi}[Dihedral Group $D_n$]
\nomenclature{Dihedral Group $D_n$}{\nomrefpage.}
    Dihedral group $D_n$ is the symmetric group of regular $n$-gons.
\end{defi}

\paragraph{Conjugacy Class}
There is some important difference between the case when $n$ is odd
and when $n$ is even. When $n$ is odd, the $2$-fold axis are all
linked with the vortices of the $n$-polygon. On the contrary, when $n$
is even, there is two different $2$-fold axis. The first $n/2$ axises
are those linked with the vortices, the second $n/2$ axises are those
cut through the opposite edges.

Note that by fact \ref{fact:20161010-srs} one can see that such
difference will note make the two classes in even $n$ the same
conjugacy class. On the contrary, by the same fact
\ref{fact:20161010-srs} one sees that those $2$-fold axises are in the
same conjugacy class when $n$ is odd.

Also denotes one $2\pi/n$ rotation around the $N$-fold axis by $T$,
then by this very same fact \ref{fact:20161010-srs}, $T$ and $T^{-1}$
lies in the same conjugacy class, related by a reflection around a
$2$-fold axis. This is obvious a self-conjugate class. (Note that,
when only $2\pi/n$ rotation is considered, they form an abelian group
and each rotation forms a conjugacy class.) Similarly for the
conjugacy class $\{T^m, T^{-m}\}$, where $m$ is an integer.

In summary, this discussion shows that the self-conjugate classes in $D_n$
are:
\begin{fact}
    When $n$ is odd (Let $n=2n'+1$): The identity class, the class of
    all $2$-fold axis. The class of $\{ T^m, T^{-m}\}$, where
    $m=1,2,\cdots, n'$. (Why $m\leq n'$, think about the case when
    $n=5,n'=2$.). In total $n'+2=\frac{n+3}{2}$ conjugacy
    classes/irreducible representations. All conjugacy classes are
    self-conjugate.
\end{fact}
\begin{fact}
    When $n$ is even (Let $n=2n'$): The identity class, the two
    classes of all $2$-fold axis, as discussed before. The class of
    $\{ T^m, T^{-m}\}$, where $m=1,2,\cdots, n'$. In total
    $n'+3=\frac{n+6}{2}$. All conjugacy classes are self-conjugate.
\end{fact}

\paragraph{Invariant Subgroups}
By the distinction between edge and vortices type of $2$-fold axises
discussed before, one can see that the difference in invariant
subgroup when $n$ is odd or even. More Specifically, when $n=5$ is odd,
$D_5$ has only $C_5$ as a notrivial subgroup. But when
$n=6$ is even, it has $C_6$ (then $C_2$ and $C_3$), two copies of
$D_3$ (each formed by edge and vortex type of $2$-fold axis).