### 2016-11-15 Lecture note - Group Theory in Physics/group theory in physics.tex
New notes for algebraic theory of representation.

% The entire content of this work (including the source code
% for TeX files and the generated PDF documents) by 
% Hongxiang Chen (nicknamed we.taper, or just Taper) is
% licensed under a 
% Creative Commons Attribution-NonCommercial-ShareAlike 4.0 
% International License (Link to the complete license text:
% http://creativecommons.org/licenses/by-nc-sa/4.0/).
\documentclass{article}

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physics-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}

\makenomenclature
% Followings are for the special character: differential "d".
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

\title{Group Theory in Physics (Course Note)}
\date{\today}
\author{Taper}

\begin{document}

\maketitle
\abstract{
This is our course note for the course about group theoy, with its
application in physics.
}
\tableofcontents
\paragraph{Information about the class}
\begin{enumerate}
    \item Final project: Combine group theory with you research.
    \item Mid-term and final exam.
    \item No homework, cause it is already graduate level.
    \item Professor Fei Ye. (Phone: 88018229, 228 in Research building 2), 
        T.A. Zhe Zhang. (110 Research building 2).
\end{enumerate}

\section{20160919}
\label{sec:20160919}
\input{sections/20160919}

\section{20160926}
\label{sec:20160926}
\input{sections/20160926}

\section{20161010}
\label{sec:20161010}
\input{sections/20161010}

\section{Skipped Two lectures}
\input{sections/Skipped-Two-Lectures}

\section{20161031}
\label{sec:20161031}
\input{sections/20161031}

\section{20161107}
\label{sec:20161107}
\input{sections/20161107}

\section{Point Group}
\label{sec:Point-Group}
\input{sections/point-group}

\section{Representationas and Semisimplicity}
\label{sec:Representationas-and-Semisimplicity}

Now I digress to read chapter 18 of the book \cite{lang-algebra}.

Let $R$ be a commutative ring and $G$ a group. Let $E$ be a
$R$-module.

For convenience, given a representation $\rho:G\to \Aut_R(E)$, let
$\sigma\ni G$, $x\in E$, we write $\sigma x$ instead of
$\rho(\sigma)x$.

An $R$-module $E$, together with a representation $\rho$, will be
called a \textbf{$G$-module}\nomenclature{$G$-module}{\nomrefpage} or
\textbf{$G$-space}\nomenclature{$G$-space}{\nomrefpage}, or also a
\textbf{$(G,R)$-module}\nomenclature{$(G,R)$-module}{\nomrefpage}.

If $E$, $F$ are $G$-modules, we define a
\textbf{$G$-homomorphism}\nomenclature{$G$-homomorphism}{\nomrefpage}
$f:E\to F$ as a $R$-linear map such that $f(\sigma x) = \sigma f(x)$
for all $x\in G$ and $\sigma\in G$.

We make $R$ into a $G$-module by making $G$ act trivially on $R$, i.e.
we have a trivial representation on $R$, itself regarded as a
$R$-module.

\section{Anchor}
\label{sec:Anchor}
\begin{thebibliography}{1}
    \bibitem{book} Zhongqi Ma, Group Theory in Physics
    \bibitem{math.se_1_lenz_vector} \href{physics.stackexchange.com/questions/18088/what-symmetry-causes-the-runge-lenz-vector-to-be-conserved}{What symmetry causes the Runge-Lenz vector to be conserved?}
    \bibitem{Ludeling} Lecture Notes for physics751: Group Theory (for
    Physicists), by C Ludeling.
    \href{http://www.th.physik.uni-bonn.de/nilles/people/luedeling/grouptheory/data/grouptheorynotes.pdf}{Link}
    \bibitem{lang-algebra} Serge Lang. Algebra. Revised 3rd. Springer.
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}