### 2016-11-15 Latex packages/latex examples.tex
Minor modification to C.V. Create a new latex example file. Begin reading lang's representaton theory for group theory in physics.

% The entire content of this work (including the source code
% for TeX files and the generated PDF documents) by 
% Hongxiang Chen (nicknamed we.taper, or just Taper) is
% licensed under a 
% Creative Commons Attribution-NonCommercial-ShareAlike 4.0 
% International License (Link to the complete license text:
% http://creativecommons.org/licenses/by-nc-sa/4.0/).
\documentclass{article}

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physicist-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}

\makenomenclature

\title{Examples}
\date{\today}
\author{Taper}

\begin{document}

\maketitle
\abstract{
Nothing here.
}
\tableofcontents
\begin{center}
    Bad diagrams
\end{center}
$$
    \begin{tikzcd}[column sep=small]
        E \ar[dr,"f"]\ar[rr,"\rho"]\ar[drr, phantom, "\#", very near
        start] & & F\ar[dl,"\rho'"] \\
        & G &  \,
    \end{tikzcd}
    \begin{tikzcd}[column sep=small]
        E \ar[dr,"f"]\ar[rr,"\rho"]\ar[drr, phantom, "\#"] & & F\ar[dl,"\rho'"] \\
        & G &  \,
    \end{tikzcd}
$$
\begin{center}
    A good diagram
\end{center}
$$
    \begin{tikzcd}[column sep=small]
        E \ar[dr,"f"]\ar[rr,"\rho"] &
        \ar[d,phantom, "\#", near start] & F\ar[dl,"\rho'"] \\
        & G &  \,
    \end{tikzcd}
$$

\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    %\bibitem{book} 
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}