### 2016-11-15 Lecture note - Group Theory in Physics/sections/s-lang.tex

\section{Representationas and Semisimplicity}
\label{sec:Representationas-and-Semisimplicity}

Now I digress to read chapter 18 of the book \cite{lang-algebra}.

Let $R$ be a commutative ring and $G$ a group. Let $E$ be a
$R$-module.

For convenience, given a representation $\rho:G\to \Aut_R(E)$, let
$\sigma\ni G$, $x\in E$, we write $\sigma x$ instead of
$\rho(\sigma)x$.

An $R$-module $E$, together with a representation $\rho$, will be
called a \textbf{$G$-module}\nomenclature{$G$-module}{\nomrefpage} or
\textbf{$G$-space}\nomenclature{$G$-space}{\nomrefpage}, or also a
\textbf{$(G,R)$-module}\nomenclature{$(G,R)$-module}{\nomrefpage}.

If $E$, $F$ are $G$-modules, we define a
\textbf{$G$-homomorphism}\nomenclature{$G$-homomorphism}{\nomrefpage}
$f:E\to F$ as a $R$-linear map such that $f(\sigma x) = \sigma f(x)$
for all $x\in G$ and $\sigma\in G$.

\begin{equation}
    \begin{tikzcd}[column sep=small]
        E \ar[rr,"f"] &
        \ar[d,phantom, "\#", near start] & F \\
        & G\ar[ul,"\rho"]\ar[ur,"\rho'"] &  \,
    \end{tikzcd}
\end{equation}

We make $R$ into a $G$-module by making $G$ act trivially on $R$, i.e.
we have a trivial representation on $R$, itself regarded as a
$R$-module.

\begin{key}
Now the we discuss a systematically way to construct a new
representation from a given one. More specifically, we want to make
$\Hom_R(E,F)$ into a $G$-module, i.e. construct a representation from
$G$ to $\Hom_R(E,F)$.

First, we define the action of $G$ on $\Hom_R(E,F)$ by (let $f\in
\Hom_R(E,F)$, $f: E\to F$).
\begin{equation}
    ([\sigma] f) (x) = \sigma (f(\sigma^{-1}x))
\end{equation}
We verify that this is an operation/action of $G$ on $\Hom_R(E,F)$:
\begin{align*}
    ([e] f) (x) &= e (f(e^{-1}x)) = f(x) \\
    ([a] [b] f )(x) &= [a] bf(b^{-1}x) = ab f(b^{-1}a^{1}x) =
    abf((ab)^{-1}x) \\
    &= ([ab] f)(x)
\end{align*}
\end{key}

\begin{fact}
\label{fact:g-hom-sg-g}
For convenience we write $([\sigma]f)(x)$ simply as $[\sigma]f(x)$.
Also, 
\begin{align*}
    f(\sigma x)=\sigma f(x) \Leftrightarrow [\sigma]f(x) = \sigma
    f(\sigma^{-1}x) = \sigma\sigma^{-1}f(x) = f(x)
\end{align*}
So $f$ is a $G$-homomorphism if and only if $[\sigma]f=f$ for all
$\sigma\in G$.
\end{fact}
\begin{remark}
    This is the first time I understand why $\sigma^{-1}$ is so
    important. Without it, we don't have the beautiful fact above.
\end{remark}
\begin{defi}[Dual representation $\rho^\vee$]
\nomenclature{Dual representation $\rho^\vee$}{\nomrefpage.}
When $F=R$ (so with trivial action), then $\Hom_R(E,F)=E^\vee$ is the
dual module. If we have $\rho:G\to \Aut_R(E)$, a representation of $G$
on $E$, then the action we just defined gives the dual
representation (also called \nomen{contragredient} in the literature).
\end{defi}
More specifically, $\sigma$ acts trivially on $R$, so
\begin{align}
    \rho^\vee: & G\to \Aut_R(E^\vee)  \\
               & \sigma \mapsto [\sigma] f = f(\sigma^{-1}x) \nonumber
\end{align}

Suppose now that the modules $E$,$F$ are free and finite dimensional
over $R$ (recall that a module is free when it has a basis). Let
$\rho$ be a representation of $G$ on $E$, $M$ be the matrix of
$\rho(\sigma)$ with respect to a basis, $M^\vee$ be the matrix of
$\rho^\vee (\sigma)$ with respect to the dual basis. Then
\begin{align*}
    [\sigma] f: x &\mapsto f(\sigma^{-1} x) \\
    x^i &\mapsto \sum_{ij} f_i(M^{-1})^i_j x^j \\
    \Longrightarrow [\sigma]: f &\mapsto f\sigma^{-1} \\
    f_j &\mapsto \sum_i f_i(M^{-1})^i_j \\
    \left(\begin{array}{c}
        f_1 \\
        f_2 \\
        \cdots 
\end{array}\right) &\mapsto (M^{-1})^T  \left(\begin{array}{c}
        f_1 \\
        f_2 \\
        \cdots 
\end{array}\right) 
\end{align*}
Hence
\begin{equation}
    M^\vee = (M^{-1})^T
\end{equation}

Next we discuss the tensor product instead of Hom. Let $E$,$E'$ be
$(G,R)$-modules. We form their tensor product $E\otimes E'$, taken over
$R$. then we define an action of $G$ on $E\otimes E'$ by (let
$\sigma\in G$):
\begin{equation}
    \sigma (x\otimes x') := \sigma x \otimes \sigma x'
\end{equation}

\begin{fact}
\label{fact:Ev-F--Hom-G}
Suppose that $E$, $F$ are free and finite dimensional over $R$. Then
the $R$-isomorphism
\begin{align}
    E^\vee \otimes F &\overset{g}{\cong} \Hom_R(E,F) \\
    f_i \otimes x^j &\overset{g}{\cong} A^j_i = (f_i x^j)
    \label{eq:fx-g-aji}
\end{align}
is a $G$-isomorphism.
\end{fact}
\begin{proof} 
In this proof, we will use Einstein summation
convention. We see that
\begin{align}
    \sigma (f_i\otimes x^j) = (\sigma^\vee f_i )\otimes (\sigma x^j) =
    f_k (M_E^{-1})^k_i \otimes (M_F)^j_l x^l 
\end{align}
So
\begin{align}
    g(\sigma (f_i\otimes x^j)) = g(f_k (M_E^{-1})^k_i \otimes 
    (M_F)^j_l x^l) = (M^{-1}_E)^k_i A_k^l (M_F)^j_l
\end{align}
Meanwhile, for $h\in \Hom_R(E,F)$,
\begin{align}
    [\sigma] h: x &\mapsto \sigma h(\sigma^{-1} x) \\
    x^i &\mapsto \sigma h^i_k (M_E^{-1})^k_j x^j \\
    &= (M_F)^i_l h^l_k (M_E^{-1})^k_j x^j
\end{align}
So $G$ acts on $h\in \Hom_R(E,F)$ is
\begin{align}
    [\sigma]: h^i_j &\mapsto  (M_F)^i_l h^l_k (M_E^{-1})^k_j
\end{align}
So, consider $g(f_i\otimes x^j)\in \Hom_R(E,F)$, we have
\begin{align*}
    [\sigma] [g (f_i\otimes x^j)] = [\sigma] A^j_i = (M_F)^j_l A^l_k
    (M_E^{-1})^k_i
\end{align*}
Hence
\begin{equation}
    \label{eq:g-iso-sggs}
    g(\sigma (f_i\otimes x^j )) = [\sigma] [g(f_i\otimes x^j)]
\end{equation}
\end{proof}

Whether $E$ is free or not, we define the \nomen{$G$-invariant
submodule of $E$} to be \nomen{$\operatorname{inv}_G(E)$}$=
R$-submodule of elements $x\in E$ such that $\sigma x=x$ for all
$\sigma\in G$. If $E,F$ are free, we have an $R$-isomorphism
\begin{equation}
    \operatorname{inv}_G (E^\vee\otimes F) \cong \Hom_G(E,F)
\end{equation}
\begin{proof}
    I give only a finite dimensional proof (for no proof is provided
    in Lang's book \cite{lang-algebra}. In finite dimensional case,
    this fact is closely related to the fact \ref{fact:Ev-F--Hom-G}:
    \begin{equation*}
        E^\vee \otimes F \overset{g}{\cong} \Hom_R(E,F)
    \end{equation*}

    Let me use the notation used before. The isomorphism is actually
    the $g$ defnied for equation \ref{eq:fx-g-aji}.

    On one hand, for $f_i\otimes x^j \in \operatorname{inv}_G
    (E^\vee\otimes F)$, denote $g(f_i\otimes x^j)$ by $A^j_i$. Then
    $g(\sigma (f_i\otimes x^j)) = g(f_i\otimes x^j) = A^j_i$ for any
    $\sigma\in G$. By equation \ref{eq:g-iso-sggs}, we have
    $[\sigma]A^j_i = A^j_i$, hence by fact \ref{fact:g-hom-sg-g}
    (notice that $A^j_i\in \Hom_R(E,F)$), $A^j_i\in \Hom_G(E,F)$.  The
    converse is similar.

    I think that for infinite dimensional cases, we might replace
    $\sum$ with $\int$ and use the same logic here.
\end{proof}