### 2016-11-21 Notes of various papers/Topological Transition in a Non-Hermitian Quantum Walk/notes.tex
Initiate new note about the 1d non-Hermitian quantum walk.

% The entire content of this work (including the source code
% for TeX files and the generated PDF documents) by 
% Hongxiang Chen (nicknamed we.taper, or just Taper) is
% licensed under a 
% Creative Commons Attribution-NonCommercial-ShareAlike 4.0 
% International License (Link to the complete license text:
% http://creativecommons.org/licenses/by-nc-sa/4.0/).
\documentclass{article}

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physicist-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}

\makenomenclature % For an index of symbols.

% Show the keys for labels, replace options with "final" when done
% with editing.
\usepackage[draft,notref]{showkeys}

\title{Notes of Topological Transition in a Non-Hermitian Quantum Walk}
\date{\today}
\author{Taper}

\begin{document}

\maketitle
\abstract{
As suggested by the title.
}
\tableofcontents
\section{General}
\label{sec:General}
This paper discusses in general a model charaterised by the follwoing
picture
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/setup.PNG}
    \caption{Setup}
\end{figure}
The equation of motion is
\begin{align}
    i\hbar \dot{\psi}^L_m = \varepsilon_L\psi^L_m + v\psi^R_m
    +v'\psi^R_{m+1} \\
    i\hbar \dot{\psi}^R_m = \tilde{\varepsilon}_R\psi^R_m + v\psi^L_m
    +v'\psi^L_{m-1}
\end{align}
\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    \bibitem{1dwalk} Topological Transition in a Non-Hermitian Quantum
    Walk,
    \href{https://arxiv.org/ct?url=http%3A%2F%2Fdx.doi.org%2F10%252E1103%2FPhysRevLett%252E102%252E065703&v=f6968ad6}{arXiv}
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}