### 2016-12-11 Basic Topology Notes/sections/ch6-Triangulation.tex
[major] new note for topology: simplicial approximation theorem

\subsection{Section 6.1 Triangulating Spaces}
\label{sec:Triangulating-Spaces}

He gave to examples first,
\begin{ex}
    A trivial example:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/ch6-scanned-notes-1/ex1.pdf}
    \end{figure}
\end{ex}
\begin{ex}
    A non trivial one. That is, the M\"obius Strip also has the
    homotopy type of a ciclr $S^1$.
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.8\linewidth]{pics/ch6-scanned-notes-1/ex2.pdf}
    \end{figure}
\end{ex}
But for complicated spaces, its homotopy type will not be so easy to
calculate in with simple imagination. Here we will introduce a new
technique to actually calculate the homotopy type of topological
spaces.

The technique is called Triangulation. The most visual example comes
from the computer vision technology (though I cannot find a picture by
directly Googling triangulation). The idea we wants to stress is that
the Triangulation is like use small patches of triangles to patch and
cover the surface of 3D smooth objects. Increase the overall number of
patches and make each patch get smaller. In the limit of this process
one might get to recover the original image.

\begin{defi}[Simplex of $\dim k$]
\nomenclature{Simplex of $\dim k$}{\nomrefpage.}
    The stanford simplex of $\dim k$ in $R^{k+1}$ is defined as
    follows. Let $v_i=(0,\cdots,0,1,0,\cdots 0)$, where $1$ is in $i+1$
    coordinates. That is:
    $$ v_0=(1,0,\cdots)$$
    $$ v_1=(0,1,0\cdots)$$
    $$\cdots$$
    $$ v_0=(0,\cdots,0,1)$$
    Then the simplex is the smallest convex set containing
    $\{v_0,\cdots,v_k\}$ in $\R^{k+1}$. It is also called a
    \nomen{$k$-simplex}.
\end{defi}
\begin{ex}
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.8\linewidth]{pics/ch6-scanned-notes-1/ex3.pdf}
    \end{figure}
\end{ex}
\begin{fact}
    Any point in a $k$-simplex is of the form:
    \begin{equation}
        \lambda_0 v_0 + \lambda_1 v_1 + \cdots + \lambda_k v_k
    \end{equation}
    where each $\lambda_i\geq 0$, and $\sum_{i=0}^{k} \lambda_k=1$.
\end{fact}
\begin{ex}
    There is a special point in the $k$-simplex:
    \begin{equation}
        \frac{1}{k+1} \sum_{i=0}^k v_i
    \end{equation}
    When $k=2$, this is just the usual center of gravity of triangle.
\end{ex}

We want to study those space which is the union of a finite collection
of simplexes which fit together nicely in some Euclidean space. These
are called "triangulable spaces". More specifically:
\begin{defi}[faces]
\nomenclature{faces}{\nomrefpage.}
    If $A$ and $B$ are simplexes, and if the vertices of $B$ form a
    subset of vertices of $A$. Then $B$ is called a face of $A$,
    written as \nomen{$B<A$}.
\end{defi}
\begin{defi}[Simplicial Complex]
\nomenclature{Simplicial Complex}{\nomrefpage.}
    A \textbf{finite} collection of simplexes in some Euclidean space
    in $\R^n$ is called a simplicial complex, if
    \begin{enumerate}
        \item whenever a simplex lies in this collection, then so does
            its faces
        \item whenever two simplexes in this collection intersect,
            their intersection is a common face of these two
            simplexes.
    \end{enumerate}
\end{defi}
\begin{ex}
    % ex 4
\end{ex}
\begin{defi}[Topology on simplicial complex]
\nomenclature{Topology on simplicial complex}{\nomrefpage.}
    The topology on a simplicial complex is given by the subspace
    topology of Euclidean space. Let $K$ be a simplicial complex, we
    denote \nomen{$|K|$} as the topology of $K$. This $|K|$ is called
    a \nomen{polyhedron}.
\end{defi}

\begin{defi}[Triangulation and Triangulable]
\nomenclature{Triangulation and Triangulable}{\nomrefpage.}
    A triangulation on a topological space $X$ consists of a
    simplicial complex $K$, and a homeomorphism $h$:
    $$ h:|K|\to X$$
    A space is called triangulable if such simplicial complex $|K|$
    and homeomorphism $h$ exists.
\end{defi}
\begin{fact}
    If $X$ is triangulable, then $X$ is compact and can be made into a
    metric space. (Notice that the simplicial complex is a finite
    collection.)
\end{fact}
\begin{remark}
    Trigulation is in general not unique. For example:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/ch6-scanned-notes-1/rmk2}
    \end{figure}
\end{remark}
\begin{ex}
    Here is a triangulation on the torus:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/ch6-scanned-notes-1/p5.pdf}
    \end{figure}
    Notice that the following diagram is not a triangulation of torus:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.8\linewidth]{pics/ch6-scanned-notes-1/p52.pdf}
    \end{figure}
\end{ex}
\begin{lemma}
    Let $K$ be a simplicial complex in $\R^{n}$, then
    \begin{enumerate}
        \item $|K|$ is compact.
        \item Each point of $|K|$ lies in the interior of exactly one
            simplex of $K$.
        \item If we take the simplexes of $K$ seperately, and give
            their union the identification topology, we obtain $|K|$.
            Thus we have two equivalent ways to view the topology on
            $K$.
        \item If $|K|$ is connected, then $|K|$ is also
            path-connected.
    \end{enumerate}
\end{lemma}
\begin{proof}
\begin{enumerate}
    \item Obviously.
    \item Since $K$ is connected by simplexes, a point $p$ must
        obviously be contained in some simplex. Also, since on each
        simplex $p$ must be on the interior of it, or interior of its
        components. We only need to prove that the containing simplex
        is unique. Suppose $p$ is the interior of simplexes $A$ and
        $B$. Notice that their intersection must be their faces. The
        only face of $A$ or $B$ containing interior points is $A$ or
        $B$ itself. Hence $A=B$.
    \item By definition of identification topology, we can see that in
        the identified space, a subset $C\subset |K|$, $C$ is closed
        $\Leftrightarrow$ $C\cap |A|$ is closed in $|A|$ for all
        simplexes $A\in K$. The rest of the proof is obvious.
    \item We need only to show that $|K|$ is locally path-connected.
        But any point $p\in K$, $p\in$ interior of some $A$ where
        $A\in K$. That is, we can find a neighbourhood $U$ of $p$, and
        $U$ is contained in $A$ ($U\subset A$). (More specifically, we
        can find $\varepsilon$ such that 
        $B_{\varepsilon}(p)\cup |K| = B_\varepsilon(p)\cup |A|$)
        But a simplex $A$ is obviously locally path-connected. Hence
        $U$ (or $B_\varepsilon(p)$) is path connected, hence $|K|$ is
        path-connected.
\end{enumerate}
\end{proof}

\subsection{Section 6.2 Barycentric Division}
\label{sec:Barycentric-Division}
Now we begin to make the triangulation more and more detailed. We
start from a simplicial complex $K$ and constructing a new simplicial
complex $K_1$, such that
\begin{enumerate}
    \item $|K|=|K_1|$, homeomorphically.
    \item diameter $|K_1|<$ diameter $|K|$.
\end{enumerate}
The general construction method is called Barycentric division,
visualized as:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/ch6-scanned-notes-1/8.pdf}
\end{figure}
We first define a concept
\begin{defi}[$\hat{A}$ Bary center]
\nomenclature{$\hat{A}$ Bary center}{\nomrefpage.}
Let $A$ be a simplex, then the bery center of $A$, denoted $\hat{A}$,
is the point:
\begin{equation}
    \hat{A}:= \frac{1}{k+1}(v_0+\cdots+v_k)
\end{equation}
\end{defi}
Then, the general process is:
The new space $K_1$ is such that
\begin{enumerate}
    \item The vertices of $K_1$ are bary centers of \textbf{ALL}
        simplexes of $K$. Note the the bary center of a $0$-simplex
        is just itself. So in general, $K_1$ has more vertices of
        $K$.
    \item A collection $\hat{A}_0,\cdots,\hat{A}_k$ of such bary
        centers will form a vertex of a $k$-simplex in $K_1$ if and
        only if there is a permutation $\sigma$ of $\{0,1,\cdots,k\}$
        such that
        \begin{equation}
            A_{\sigma{0}}< A_{\sigma{1}} < \cdots < A_{\sigma{k}}
        \end{equation}
\end{enumerate}
\begin{remark}
    The second point above tries to do the following
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.9\linewidth]{pics/ch6-scanned-notes-1/p9.pdf}
    \end{figure}
\end{remark}

Now we give a precise description of how this Bary center division
will work. We first describe precisely two characteristics of a
simplicial complex.

\begin{defi}[dimension of a simplicial complex]
\nomenclature{dimension of a simplicial complex}{\nomrefpage.}
    The dimension of a simplicial complex $K$ is the maximum of the
    dimensions of all its simplexes.
\end{defi}
\begin{defi}[mesh $\mu(K)$]
\nomenclature{mesh $\mu(K)$}{\nomrefpage.}
    The mesh of a simplicial complex $K$ is the maximum of the
    diameters of all its simplexes.
\end{defi}

As we expected, the mesh (which carries the intuition of the smallness
of one simplicial complex) of a simplicial complex will decrease. The
following lemma confirmed this.

\begin{lemma}
    The sollection of simplexes described above forms a simplicial
    complex. If it is denoted by $K^1$, and is called the first
    barycentric subdivision of $K$, then $K^1$ has the following
    properties:
    \begin{enumerate}
        \item each simplex of $K^1$ is contained in a simplex of $K$;
        \item Their topology are the same: $|K^1|=|K|$
        \item If the dimension of $K$ is $n$, then
            \begin{equation}
                \mu(K^1)\leq \frac{n}{n+1}\mu(K)
            \end{equation}
    \end{enumerate}
\end{lemma}
\begin{proof}
    The proof can be found in page 126 to 127 of \cite{book}. But the
    proof of the last point is not so clear and is illustrated here:
\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{pics/ch6-notes-2/explain1.pdf}
\end{figure}
\end{proof}
\begin{remark}
    Notice that the number $\frac{n}{n+1}<1$, hence after an infinite
    number of Barycentric subdivision, the mesh (averaged diameter) of
    a simplicial complex $K^m$ will goes to zero!
\end{remark}
\subsection{Section 6.3 Simplicial approximation}
\label{sec:Simplicial-approximation}

The purpose here is to approximate any continuous map $f:A\to B$ with
linear map, in a way that is compatible with the triangulations on
both $A$ and $B$.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.4\linewidth]{pics/ch6-notes-2/exp2.png}
\end{figure}
\begin{defi}[Simplicial map]
\nomenclature{Simplicial map}{\nomrefpage.}
    Let $K$ and $L$ be simplicial complexes. A continuous map $s:K\to
    L$ is called simplicial, if it takes simplexes of $K$
    \textit{linearly} to simplexes of $L$. Specifically, being
    \textit{linear} means that:

    For $\forall A\in K$, we have $B=s(A)\in L$. Also, if
    $A=\{v_0,\cdots,v_k\}$, then $B=\{s(v_0),\cdots,s(v_k)\}$.
    Lastly, if $x=\sum_{i=0}^k \lambda_i v_i$ is a a point in $A$, with
    $\lambda_i\geq 0$ and $\sum_{i=0}^k\lambda_i=1$, then
    $$s(x) = \sum_{i=0}^k \lambda_i s(v_i)$$

    In a word, $s$ maps simplexes to simplexes, vertices, to vertices,
    and points linearly to points.
\end{defi}
\begin{remark}
    It is easy to see that the a simplicial map is uniquely determined
    by its value on all the vertices. By this, there can only be a
    finite number of such kind of maps, since there are only a finite
    number of vertices.
\end{remark}

Now we want to find such linear maps to approximate a give function
$f:|K|\to |L|$. Given a point $x\in |K|$, we define the
\nomen{carrier of $f(x)$} to be the unique simplex of $L$ that contain
$f(x)$ as its interior point.

\begin{defi}[Simplicial Approximation]
\nomenclature{Simplicial Approximation}{\nomrefpage.}
    A simplicial map $s:|K|\to |L|$ is called a simplicial approximation of
    $f:|K|\to |L|$ if and only if $s(x)\in$ carrier of $f(x)$,
    $\forall x\in |K|$.
\end{defi}
\begin{ex}
    Trivial examples:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/ch6-notes-2/ex3.png}
\end{figure}
\end{ex}
\begin{ex}
    Now we show that when simplicial approximation does not exits. The
    set up is as follows:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.8\linewidth]{pics/ch6-notes-2/ex4.png}
    \end{figure}
    Hence, there is no simplicial approximation of $f$. But, it will
    be shown below that, if we divide $K$ small enough
    ($K^1,K^2,\cdots$), we will eventually be able to find a
    simplicial approximation of $f$ in some $K^n$.
\end{ex}

\begin{thm}[Simplicial approximation theorem]
    \label{thm:SAT}
    Let $f:|K|\to |L|$ be a continuous map between polyhedras formed
    by simplicial complexes $|K|$ and $|L|$. If $m$ is chosen large
    enough, then there is a simplicial approximation $s:|K^m|\to|L|$
    to $f:|K^m|\to|L|$.
\end{thm}
The proof will use a lemma. Define the \nomen{open star} $v$ of a
simplicial complex $K$, as the union of the interiors of those
simplexes of $K$ having $v$ as a vertex. This union is denoted by
\nomen{$\operatorname{star}(v,K)$}. An example is:
\begin{figure}[H]
    \centering
    \includegraphics[width=1.0\linewidth]{pics/ch6-notes-2/ex5.png}
    \caption{Example of star}
\end{figure}
So it really feels like a star spreading out from $v$. We have

\begin{lemma}
    Vertices $v_0,v_1,\cdots,v_k$ of a simplicial complex $K$ are the
    vertices of a simplex of $K$ if and only if the intersection of
    their open stars is nonempty, i.e. 
    $$\cup_0^k \operatorname{star}(v_i,K)\neq\emptyset$$
\end{lemma}
\begin{proof}
    The proof can be found in page 130 of \cite{book}. It is quite
    trivial.
\end{proof}
Then we have the \textbf{proof of theorem \ref{thm:SAT}}:
\begin{proof}
    The proof is found in page 130 to 131 of \cite{book}. The idea is
    that if for each vertex $u$ of $K$, we can find a vertex $v$ of
    $L$ such that
    \begin{equation}
        f(\operatorname{star}(u,K))\subseteq\operatorname{star}(v,L)
        \label{eq:proof-temp} \tag{*} 
    \end{equation}
    Then the simplicial map $s(u)=v$ will be easily verified to be a
    simplicial approximation of $f$.

    To make such (*) happen, we naturally have to divide $K^m$ small
    enough. We can achieve such small division is guaranteed by the
    Lebesgue lemma \ref{lemma:lebesgue-lemma}, since each subdivision
    makes the averaged diameter smaller.
\end{proof}