### 2017-02-17 Lecture note - Group Theory in Physics/group theory in physics.tex
[minor] Uploading some missing notes about Group theory in Physics.

% The entire content of this work (including the source code
% for TeX files and the generated PDF documents) by 
% Hongxiang Chen (nicknamed we.taper, or just Taper) is
% licensed under a 
% Creative Commons Attribution-NonCommercial-ShareAlike 4.0 
% International License (Link to the complete license text:
% http://creativecommons.org/licenses/by-nc-sa/4.0/).
\documentclass{article}

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physics-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}

\makenomenclature
% Followings are for the special character: differential "d".
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

% Show the keys for labels, replace options with "final" when done
% with editing.
\usepackage[draft,notref]{showkeys}

% Resolve pdf version warning
\pdfoptionpdfminorversion=6

\title{Group Theory in Physics (Course Note)}
\date{\today}
\author{Taper}

\begin{document}

\maketitle
\abstract{
This is our course note for the course about group theoy, with its
application in physics.
}
\tableofcontents
\paragraph{Information about the class}
\begin{enumerate}
    \item Final project: Combine group theory with you research.
    \item Mid-term and final exam.
    \item No homework, cause it is already graduate level.
    \item Professor Fei Ye. (Phone: 88018229, 228 in Research building 2), 
        T.A. Zhe Zhang. (110 Research building 2).
\end{enumerate}

\section{20160919}
\label{sec:20160919}
\input{sections/20160919}

\section{20160926}
\label{sec:20160926}
\input{sections/20160926}

\section{20161010}
\label{sec:20161010}
\input{sections/20161010}

\section{Skipped Two lectures}
\input{sections/Skipped-Two-Lectures}

\section{20161031}
\label{sec:20161031}
\input{sections/20161031}

\section{20161107}
\label{sec:20161107}
\input{sections/20161107}

\section{Point Group}
\label{sec:Point-Group}
\input{sections/point-group}

\section{20161121}
\input{sections/20161121}
\section{20161128}
\label{sec:20161128}
After some work, I realize that it is impractical to follow the
teacher's class with notes. Therefore I will write here just some
important formulae, as a preparation for exams.

Pauli matrices
\begin{align}
    &\sigma_1 =\begin{pmatrix}
        0 & 1\\ 1 & 0
    \end{pmatrix}\\
    &\sigma_2 = \begin{pmatrix}
        0& -i \\ i & 0
    \end{pmatrix} \\
    &\sigma_3 = \begin{pmatrix}
    1 & 0 \\ 0 & -1
    \end{pmatrix}\\
    &\sigma_a \sigma_b = \delta_{ab} \id + i\sum_c
    \varepsilon_{abc}\sigma_c \\
    &\Tr{\sigma_a} =0 \\
    &\Tr{\sigma_a \sigma_b} =2\delta_{ab} \\
    &\exp{-i\omega \sigma_2} =\id \cos{\omega}-i\sigma_2\sin{\omega}
    = \begin{pmatrix}
        \cos{\omega} & -\sin{\omega} \\
        \sin{\omega} & \cos{\omega}
    \end{pmatrix} \\
    &R(\vec{e_3},\omega)= \exp{-i\omega T_3}= \begin{pmatrix}
        \cos{\omega} & -\sin{\omega} & 0 \\
        \sin{\omega} & \cos{\omega} & 0 \\
        0 & 0 & 1
    \end{pmatrix},\quad T_3 = \begin{pmatrix}
        0 & -i & 0 \\
        i & 0 & 0 \\
        0 & 0 & 0
    \end{pmatrix} \\
    & R(\vec{e}_1,\omega) = \exp{-i\omega T_1} = \begin{pmatrix}
        1 & 0 & 0 \\
        0 & \cos{\omega} & -\sin{\omega} \\
        0 & \sin{\omega} & \cos{\omega} 
    \end{pmatrix},\, T_1 = \begin{pmatrix}
        0 & 0 & 0 \\
        0 & 0 & -i \\
        0 & i & 0
    \end{pmatrix}\\
    & R(\vec{e}_2,\omega) = \exp{-i\omega T_2} = \begin{pmatrix}
        \cos{\omega} & 0 & \sin{\omega} \\
        0 & 1 & 0 \\
    -\sin{\omega} & 0 & \cos{\omega} 
    \end{pmatrix},\, T_2 = \begin{pmatrix}
        0 & 0 & i \\
        0 & 0 & 0 \\
        -i & 0 & 0
    \end{pmatrix}\\
    & (T_a)_{bc}= -i\varepsilon_{abc} \\
    & S(\phi,\theta) = \begin{pmatrix}
    \cos{\phi}\cos{\theta} & -\sin{\phi} & \cos{\phi}\sin{\theta} \\
    \sin{\phi}\cos{\theta} & \cos{\phi} & \sin{\phi}\sin{\theta} \\
    -\sin{\theta} & 0 & \cos{\theta}
    \end{pmatrix} \\
    & S(\phi,\theta) \begin{pmatrix}
        0 \\ 0 \\ 1
    \end{pmatrix} = \begin{pmatrix}
        \cos{\phi}\sin{\theta} \\ \sin{\phi}\sin{\theta} \\ \cos{\theta}
    \end{pmatrix} = \begin{pmatrix}
        n_1 \\ n_2 \\ n_3
    \end{pmatrix}\\
    & R(\vec{n},\omega) =
        S(\phi,\theta)R(\vec{e_3},\omega)S(\phi,\theta)^{-1}
        =\exp{-i\omega\vec{n}\vdot\vec{T}}
        = \exp{-i\sum_a \omega_a T_a} \\
    & R(\vec{n},\omega) = R(-\vec{n},2\pi-\omega) \\
    & S T_3 S^{-1} = \vec{n}\vdot\vec{T}
\end{align}
\begin{fact}
    The character $\chi$ of $SO(3)$ is a function of $\omega$ only. In
    fact:
    \begin{equation}
        \chi(R(\vec{n},\omega)) = 1+2\cos{\omega}
    \end{equation}
\end{fact}
\begin{enumerate}
    \item $R\in G$, $\# G=g$

        $ R(1):=R(r_1,r_2,\cdot,r_g)$
    \item $R(r)S(s)=T(t)$,

        $f(r,s)=t$, $f:(Space)\otimes(Space)\to (Space)$.

        $t_j = f_j(r_1,\cdot,r_g,s_1,\cdot,s_g)=f_j(r,s)$
\end{enumerate}

\begin{align}
    & P_R\psi(x)=\psi(R^{-1}x) \\
    & I^{(0)}_j = -i\sum_a 
        \eval{\frac{\partial (Ax)_a}{\partial \alpha_j}}_{\alpha=0}
        \frac{\partial }{\partial x_a} \\
    & P_A\psi(x) = \psi(x)-i\sum_{j=1}^g \alpha_j I^{(0)}_j\psi(x)
\end{align}
\section{Anchor}
\label{sec:Anchor}
\begin{thebibliography}{1}
    \bibitem{book} Zhongqi Ma, Group Theory in Physics
    \bibitem{math.se_1_lenz_vector} \href{physics.stackexchange.com/questions/18088/what-symmetry-causes-the-runge-lenz-vector-to-be-conserved}{What symmetry causes the Runge-Lenz vector to be conserved?}
    \bibitem{Ludeling} Lecture Notes for physics751: Group Theory (for
    Physicists), by C Ludeling.
    \href{http://www.th.physik.uni-bonn.de/nilles/people/luedeling/grouptheory/data/grouptheorynotes.pdf}{Link}
    \bibitem{lang-algebra} Serge Lang. Algebra. Revised 3rd. Springer.
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}