### 2017-03-25 Notes of Bernevig's Book/sections/chapter2-berry-phase.tex

\subsection{Deriving Berry Phase}
\label{sec:Deriving-Berry-Phase}
From page 1 to page 2 (equation 2.7), one derives the expression for
the phase $e^{i\gamma_m}$ using instaneous energy eigenstate for
$E_m$:
\begin{align}
    & H(\vec{R})\ket{n(\vec{R})} = E_m(\vec{R})\ket{n(\vec{R})} \\
    & \gamma_m = i\int_0^t
    \braket{
        m\left(\vec{R}(t')\right)|\frac{\partial }{\partial t}
        m\left(\vec{R}(t')\right) \dd t' }
\end{align}
Then I writes about different ways to get the $\gamma_m$:
\begin{align}
    \gamma_m = i\int_{\text{curve}}\braket{m|\nabla_{\vec{R}}\, m }
    \dd \vec{R}
    = \int_{\text{curve}} \vec{A}_n \vdot\dd \vec{R}
\end{align}
(equation 2.8) 
where one defines
\begin{defi}[Berry Connection, Berry Vector Potential $\vec{A}$]
\nomenclature{Berry Connection, Berry Vector Potential
$\vec{A}$}{\nomrefpage.}
    \begin{align}
        \vec{A}_n \equiv i \braket{n|\nabla_{\vec{R}}\, n }
    \end{align}
\end{defi}
Then I proves several facts:
\begin{fact}
    \label{fact:gamma_is_real}
        $\gamma_n$ is real
\end{fact}
By virtue of fact \ref{fact:gamma_is_real}, we have
\begin{align}
    \gamma_n = -\Im \int_\text{curve} \braket{ n|\nabla_{\vec{R}}\,n}
    \vdot \dd \vec{R} 
\end{align}
\begin{fact}
    Berry connection $\vec{A}_n$ is gauge-dependent.
    The dependence is: If 
    $$\ket{n}\to \ket{n'}= e^{i\xi(\vec{R})}\ket{n}$$
    then
    \begin{align}
        \vec{A}_n \to \vec{A}_n - \nabla_{\vec{R}}\,\xi(\vec{R})
    \end{align}
\end{fact}
Therefore we have
\begin{fact}
    $\gamma_n$ is gauge-dependent, unless the path transverses a
    closed loop.
    Since 
    $$ \gamma_n \to \gamma_n - \left(\xi(\vec{R}(T))-
    \xi(\vec{R}(0)) \right) $$
    It is unchanged unless the integration curve is a closed loop. In
    which case
    $$\xi(\vec{R}(T)) - \xi(\vec{R}(0)) = 2\pi m
    \overset{\text{mod }2\pi}{=} 0 $$
\end{fact}
\begin{ex}
    There is a simple example on page 3 to show that the Berry phase
    can be actually detected.
\end{ex}

When the parameter space is $\mathbb{R}^3$, we have a simpler
expression for berry phase. It is derived on page 4 (equation 2.12)
that
\begin{align}
    \gamma_n = -\Im \oint \braket{\grad_{\vb{R}}
    n|\cross|\grad_{\vb{R}} n} \vdot \dd \vb{s}
    = -\Im \oint \mathcal{A}^n \vdot \dd\vb{s}
\end{align}
where we have defined:
\begin{defi}[Berry Curvature $\mathcal{A}^n$]
\nomenclature{Berry Curvature $\mathcal{A}^n$}{\nomrefpage.}
    \begin{align}
        \mathcal{A}^n = \braket{\grad_{\vb{R}} n|\cross|\grad_{\vb{R}} n}
    \end{align}
    In components (repeated index automatically summed):
    \begin{align}
        \mathcal{A}_i^n = \varepsilon_{ijk} \braket{\partial_j n |
        \partial_k n}
    \end{align}
\end{defi}
By analogy with the theory of electromagnetic fields, this is also
called Berry field, dentoed by $F_{jk}$ in Bernevig's book.
\nomenclature{Berry field $F_{jk}$}{\nomrefpage}. More specifically:
$F_{jk} = \braket{\partial_j n(\vb{R})|\partial_j n(\vb{R})} - (j
    \leftrightarrow k) $.

\subsection{Gauge-independent calculation of Berry Curvature}
\label{sec:Gauge-independent-calculation-of-Berry-Curvature}
For numerical considerations, we Bernevig gives a new way to calculate
the Berry curvature. I shows in page 5 to 6 that:
\begin{align}
    \gamma_n = -\Im \oint \dd\vb{s} \sum_{m\neq n}
    \frac{\bra{n(\vb{R})}\grad_{\vb{R}} H \ket{m(\vb{R})}
        \cross \bra{m(\vb{R})}\grad_{\vb{R}} H \ket{n(\vb{R})} }
        {(E_n-E_m)^2}
\end{align}
There are two advantages of this formula:
\begin{enumerate}
    \item It is intrinsically gauge-independent (see page 6 for
        explanation).
    \item It is nolonger necessary to pick $\ket{n}$ to be smooth \&
        single-valued. \marginpar{Don't know why}
\end{enumerate}
There are several remarks about this formula on page 7. One important
remark is that:
\begin{fact}
    \begin{align}
        \sum_n \gamma_n = 0
    \end{align}
\end{fact}

\subsection{Degeneracy and level-crossing}
\label{sec:Degeneracy-and-level-crossing}
Page 7 has a discussion about the serious problems posed by the
degenracy points. Then we turn our attention to two-level system.

    \subsubsection{Two-level system and Berry connection
    \texorpdfstring{$\vb{A}_n$}{}}
    \label{sec:Two-level-system-and-Berry-connection}
    Here Bernevig considers a system with Hamiltonian
    \begin{equation}
        H = \epsilon(\vb{R}) I + \vb{d}(\vb{R})\vdot\vb{\sigma}
    \end{equation}

    He first calculate in a general setting and concludes that the Berry
    field is
    \begin{equation}
        \vb{V}_{-} = \frac{1}{2}\frac{\vb{d}}{d^3}
    \end{equation}
    However, I don't think it is a good calculation, explained in page 8.
    On the other hand, Bernevig also gives an alternative calculation in
    the following.

        \paragraph{Two-level system using Hamiltonian approach}
        \label{sec:Two-level-system-using-Hamiltonian-approach}
        On page 8 to 10, I gives the calculation of Berry phase of
        Spin in a varying magnetic environment. Consider the
        Hamiltonian:
        \begin{equation}
            H(\vb{B}) = \vb{B}\vdot \vb{S} 
        \end{equation}
        with
        \begin{equation}
            n = -s, -s+1,\cdots, s
        \end{equation}
        One finds:
        \begin{align}
            \gamma_n = - n\oint \frac{\vb{B}}{B^3} \dd\vb{s}
        \end{align}
        (equation 2.31)
        If the integration is taken on the surface of a sphere, then
        we have (noting that the solid angle $\dd
        \Omega=\sin(\theta)\dd\theta\dd\phi$. Use $\Omega$ to denote
        the solid angle of the integration area)
        \begin{equation}
            e^{i\gamma_n} = e^{-i n \Omega}
        \end{equation}

        There are several remarks on page 10. The important ones are
        \begin{remark}
            If we integrate in a whole sphere, and the sphere contains
            several "poles" - the degeneracy point, we have
            \begin{align}
                |\gamma_i| = 4\pi n \times \text{(number of poles
                inside)}
            \end{align}
            Similary, if the paremeter space in 2 dimensional, then
            \begin{align}
                |\gamma_i| = 2\pi n \times \text{(number of poles
                inside)}
            \end{align}
            Specifically, for electrons
            \begin{align}
                |\gamma_i| = \pi \times \text{(number of poles
                inside)}
            \end{align}
            Therefore, an electron transversing a circle will
            accumulate a phase of $e^{i\pi}=-1$.
        \end{remark}