### 2017-03-29 Notes of various papers/TR invariant topological insulators T. Hughes. Dissertation/note.tex
[minor] Update something about spin-waves in this model...

% The entire content of this work (including the source code
% for TeX files and the generated PDF documents) by 
% Hongxiang Chen (nicknamed we.taper, or just Taper) is
% licensed under a 
% Creative Commons Attribution-NonCommercial-ShareAlike 4.0 
% International License (Link to the complete license text:
% http://creativecommons.org/licenses/by-nc-sa/4.0/).
\documentclass{article}

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physicist-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}
\makenomenclature % For an index of symbols.

\title{TR Invariant T.I.}
\date{\today}
\author{Taper}

\begin{document}

\maketitle
\abstract{
    An incomplete note of dissertation by Taylor Hughes
    \cite{Hughes2009}.
}
\tableofcontents
Start with chapter 2.

Questions:
\begin{itemize}
    \item Do we have a precise definition of \textit{topological phase
        transition}?
\end{itemize}

\section{Spectrum of \texorpdfstring{$(2+1)$}-d Lattice Dirac Model}
\label{sec:2+1d-LDirac Model}
\begin{align}
    H_{LD} &= \sum_{m,n}\left\{
        i \left[ c^\dagger_{m+1,n}\sigma^x c_{m,n}
            - c^\dagger_{m,n}\sigma^x c_{m+1,n}\right]
        + i \left[ c^\dagger_{m,n+1}\sigma^y c_{m,n}
            - c^\dagger_{m,n}\sigma^y c_{m,n+1} \right]
        \right.
        \nonumber\\
        &- \left[
            c^\dagger_{m+1,n}\sigma^z c_{m,n}
            + c^\dagger_{m,n}\sigma^z c_{m+1,n}
            + c^\dagger_{m,n+1}\sigma^z c_{m,n}
            + c^\dagger_{m,n}\sigma^z c_{m,n+1}
        \right]
        \nonumber\\
        & \left. + (2-m) c^\dagger_{m,n}\sigma^z c_{m,n}
        \vphantom{xx}\frac\hbar2
    \right\}
\end{align}

Above is the lattice model (eq.2.19) of \cite{Hughes2009}. Here it
should be noted that $c_{m,n}=(c_{u,m,n}, c_{v,m,n})$ for two degrees
of freedom.

\subsection{Numerical Solution in Infinity Cylinder Geometry}
\label{sec:Numerical Solution in Infinity Cylinder Geometry}

This Hamiltonian is solved here with a infinite cylinder geometry, i.e.
the lattice is infinite in $x$ direction while being periodic in $y$
direction. Because of this special setup, the $p_x$ is still a good
quantum number. Therefore we can do a fourier expansion in $x$
direction:

\begin{equation}
    c_{m,n} = \frac{1}{\sqrt{L_x}} \sum_{p_x} e^{ip_x m} c_{p_x,n}
\end{equation}

The resulted Hamiltonian is

\begin{align}
    \tilde H_{LD} = \sum_{n,p_x}\quad &
    2\sin(p_x) c^\dagger_{p_x,n}\sigma^x c_{p_x,n}
    +i \left[ c^\dagger_{p_x,n+1}\sigma^y c_{p_x,n} - 
        c^\dagger_{p_x,n+1}\sigma^y c_{p_x,n}\right]
    \nonumber\\
    &- \left[ 2\cos(p_x) c^\dagger_{p_x,n}\sigma^z c_{p_x,n}
        c^\dagger_{p_x,n+1}\sigma^z c_{p_x,n}
        + c^\dagger_{p_x,n}\sigma^z c_{p_x,n+1}
    \right]
    \nonumber\\
    &+ (2-m) c^\dagger_{p_x,n}\sigma^z c_{p_x,n}
\end{align}

This Hamiltonian can be solved by acting it on the test wavefunction:
\begin{equation}
    \ket{\psi_{p_x}} = \sum_n \psi_{p_x,n,u} c^\dagger_{p_x,n,u} +
                        \psi_{p_x,n,v} c^\dagger_{p_x,n,v} \ket{0}
\end{equation}

Note, in choosing the test wavefunction, $u$ and $v$ could not be
seperated, because there is still interaction between the two
component in terms like $c^\dagger_{p_x,n}\sigma^x c_{p_x,n}$.
If we calculate $\tilde H_{LD}\ket{\psi_{p_x}} = E_{p_x}
\ket{\psi_{p_x}}$, we would get after careful calculation:

\begin{align}
    & \sum_n
    c^\dagger_{p_x,n} A   \psi_{p_x,n-1}
    + c^\dagger_{p_x,n} B \psi_{p_x,n} 
    + c^\dagger_{p_x,n} C \psi_{p_x,n+1}
    \nonumber\\
    &= E_{p_x} \sum_n c^\dagger_{p_x,n}\psi_{p_x,n}
\end{align}
where
\begin{align}
    c^\dagger_{p_x,n} &= \begin{pmatrix}
        c^\dagger_{p_x,n,u},  c^\dagger_{p_x,n,v}
    \end{pmatrix} \\
    A &= i\sigma^y - \sigma^z \\
    B &= 2\sin(p_x)\sigma^x - 2\cos(p_x)\sigma^z + (2-m)\sigma^z \\
    C &= -i\sigma^y - \sigma^z \\
    \psi_{p_x,n} &= \begin{pmatrix}
        \psi_{p_x,n,u} \\ \psi_{p_x,n,v}
    \end{pmatrix}
\end{align}

Suppose there is $N$ lattice in the $y$ direction. Then the periodic
boundary condition implies that $\psi_{N+1}=\psi_{n=1}$, and
$\psi_{n=0} = \psi_N$.

Therefore, the eigenvalue equation could be turned into a matrix form:
\begin{align}
    H_{\text{disc}} \psi \equiv 
    \begin{pmatrix}
        B   & C &        &   & A \\
        A   & B & C      &   & \\
            & A & B      & C & \\
            &   & \cdots &   & \\
            &   & A      & B & C \\
        C   &   &        & A & B
    \end{pmatrix}
    \begin{pmatrix}
        \psi_{p_x,1} \\
        \psi_{p_x,2} \\
        \cdots \\
        \\
        \\
        \psi_{p_x,N} \\
    \end{pmatrix} =
    E_{p_x}
    \begin{pmatrix}
        \psi_{p_x,1} \\
        \psi_{p_x,2} \\
        \cdots \\
        \\
        \\
        \psi_{p_x,N} \\
    \end{pmatrix}
\end{align}

\textbf{Note}: Numerical calculations in this section are contained in the file
"Lattice Dirac Model (2+1)-d.nb", and the file
"Dirac\_Lattice\_Model\_21\_d.m".

Let us take $N=3$ for simplicity. The eigenvalue problem is solve
using Mathematica, and the $6$ eigenvalues are:
\begin{equation}
    \begin{pmatrix}
    -\sqrt{m^2+4 m \cos (\text{px})+4}\\
    \sqrt{m^2+4 m \cos (\text{px})+4}\\
    -\sqrt{m^2+4 m \cos (\text{px})-6 m-12 \cos (\text{px})+16}\\
    -\sqrt{m^2+4 m \cos (\text{px})-6 m-12 \cos (\text{px})+16}\\
    \sqrt{m^2+4 m \cos (\text{px})-6 m-12 \cos (\text{px})+16}\\
    \sqrt{m^2+4 m \cos (\text{px})-6 m-12 \cos (\text{px})+16}
    \end{pmatrix}
\end{equation}

It is found that at $m=-2$, there is a band crossing at $p_x=0$:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.2\linewidth]{pics/N6m-2.pdf}
    \caption{The Eigenvalue plot for $m=-2$. Plotted as
    $E_{p_x}$-$p_x$ }
\end{figure}

Also, at $m=2$, there is a band crossing at $p=\pm\pi$:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/N6m2.pdf}
    \caption{The Eigenvalue plot for $m=2$. Plotted as
    $E_{p_x}$-$p_x$ }
\end{figure}

When the band crosses, there will be two eigenvectors, corresponds to
the two crossed bands, in the form of:
\begin{align}
    \psi_{p_x} &= \left(\vphantom{\frac12}\psi(p_x),1,\psi(p_x),1,\psi(p_x),1\right)^T \\
    \phi_{p_x} &= \left(\vphantom{\frac12}\phi(p_x),1,\phi(p_x),1,\phi(p_x),1\right)^T
\end{align}
where $\psi(p_x)$ and $\phi(p_x)$ are functions of $p_x$. A look into
the plot of $\psi(p_x)$ and $\phi(p_x)$ reveals that they together
provide the path way for excited particles to transfer from the lower
band to the upper band.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/upperSpinWaveMN2.pdf}
    \caption{Plot of $\psi(p_x)$ and $\phi(p_x)$ when $m=-2$}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/upperSpinWaveM2.pdf}
    \caption{Plot of $\psi(p_x)$ and $\phi(p_x)$ when $m=2$, where I
    have extended the range s.t. $p_x\in\{-\pi,3\pi\}$ to make the
    meaning clear.}
\end{figure}

Therefore, one may predict a pure spin-up wave transfering in the
point $p_x=0$ when $m=2$, and $p_x=\pm\pi$ when $m=2$.

    \subsection{Why I think the Lattice Model is wrong}

    I notice that equation (2.19) transformed according to (2.20)
    is not exactly equation (2.21), but is:
    \begin{align}
        &H =\sum_{p_x,p_y} c^\dagger_{p_x,p_y}\times
        \nonumber\\
        &
        \left[
            2\sin(p_x)\sigma^x + 2\sin(p_y)\sigma^y
            +(2-m-2\cos(p_x)-2\cos(p_y))\sigma^z
        \right] c_{p_x,p_y}
    \end{align}
    This result does not become the continuum Dirac Hamiltonian as
    $p_x,p_y$ goes to zero. Therefore, I suspect that certain
    constants should be modified so that:

    \begin{align}
        H_{LD} &= \sum_{m,n}\left\{
            \frac{i}{2} \left[ c^\dagger_{m+1,n}\sigma^x c_{m,n}
                - c^\dagger_{m,n}\sigma^x c_{m+1,n}\right]
            + \frac{i}{2} \left[ c^\dagger_{m,n+1}\sigma^y c_{m,n}
                - c^\dagger_{m,n}\sigma^y c_{m,n+1} \right]
            \right.
            \nonumber\\
            &- \frac{1}{2}\left[
                c^\dagger_{m+1,n}\sigma^z c_{m,n}
                + c^\dagger_{m,n}\sigma^z c_{m+1,n}
                + c^\dagger_{m,n+1}\sigma^z c_{m,n}
                + c^\dagger_{m,n}\sigma^z c_{m,n+1}
            \right]
            \nonumber\\
            & \left. + (2-m) c^\dagger_{m,n}\sigma^z c_{m,n}
            \vphantom{\frac12}
        \right\}
    \end{align}

    This affects the numerical analysis effectively by the replacement
    $$\sigma^i \to \frac{1}{2}\sigma^i,\quad
        (2-m) \to 2(2-m) $$

    The calculated result is similar to that in the previous section,
    except that the band crossing happens at different values of $m$.
    \footnote{For example, the eigenvalue of original and the modified equation
    (2.21) are plotted in Mathematica notebook "Eq2.21-Demo.nb". Also,
    the solution to the infinite cylinder boundary condition has again
    two band crossings, each at $(m,p_x)$ equals $(0,0)$ and
    $(2,\pm\pi)$ (for $N=3$ case).}
    So the essential point is unaltered by the difference in some
    constants. However, in the correct calculation, the crossing band
    appears at $m=0$, which represents a massless spin-$\frac{1}{2}$
    particle. I think this should have some theoretical implications.

\subsection{Calculation Note unrelated to the main-subject}

Since the paper will be focusing in points around $p_x=0$, I focused
in $m=-2$ at first. In this case, I want to find more information
about the eigenvectors.

When I looked blindly at the value $(m,p_x)=(-2,0)$, the
Mathematica gave me two eigenvectors both corresponds to the
eigenvalue $0$:
\begin{equation}
    \{0,1,0,1,0,1\},\{1,0,1,0,1,0\}
\end{equation}
It led me to believe that there are two spin waves, with made with
purely spin up waves and another of purely spin down waves. But
this is not correct.

It is found later that the matrix $H_\text{disc}$ is singular
(with determinant $0$) when $(m,p_x)=(-2,0)$. Also, a Matlab
calculation shows that the eigenvectors of the crossing bands
actually flunctuate between $\pm 1$ in a way illustrated as below:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/Eigenvector-m.eps}
\end{figure}

Also, the Mathematica solved eigenvector also demonstrate a
drastical change around $m=-2$. For example, one component, when
plotted against $p_x$ change from:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/mN3.pdf}
    \caption{$m=-3$}
\end{figure}
to
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/{mN2.5}.pdf}
    \caption{$m=-2.5$}
\end{figure}
and suddenly to
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/mN2.pdf}
    \caption{$m=-2$. There is a discontinuity at $p_x=0$}
\end{figure}
Finaly, it becomes smooth again:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/{mN1.5}.pdf}
    \caption{$m=-1.5$}
\end{figure}
The details can be explored in the Mathematica notebook.

Also, the case of $N=4$ is also calculated in Mathematica.
There are similarly two crossing happening at $(m,p_x)$ equals
$(-2,0)$ and $(2,\pm\pi)$.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/N8m2.pdf}
    \caption{$m=2$}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/N8m-2.pdf}
    \caption{$m=-2$}
\end{figure}

Surprisingly, the two bands that cross are have exactly the same
function dependence on $p_x$ and $m$ for the cases of $N=3$ and
$N=4$.

\bibliography{cite}{}
\bibliographystyle{alphaurl}

\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}