### 2017-05-16 Notes of various papers/Chiu Minimal Dirac Hamiltonian Method/chapters/chap-appA.tex

\section{Various Properties for Clifford Algebras}
\label{sec:Various Properties for Clifford Algebras}
We have used various properties of Clifford Algebras from various sources. Here
I summarized those properties for the convenience of readers.

\paragraph{Isomorphism Relations}
The best source of isomorphism between Clifford Algebras is actually the
Wikipedia page about \textit{Classifications of Clifford Algebras}. For a more
authoritative account, I am using the book \cite{VazJayme2016}. The results
used in this document are to be found in chapter 4 of that book.

\paragraph{Representation Theory on Complex Clifford Algebras}
This paper \cite{West1998} has a good discussion on the irreducible
representations of complex Clifford Algebras, in p.5, p.12. They can be
summarized as:
\begin{thm}
	Consider $\mathrm{C\ell}_{N}(\C)$. For both $N = 2n$ and $N = 2n + 1$, the
	dimension d of the irreducible representation is uniquely given by
	\begin{equation}
	d = 2^n \text{($N=2n$ or $N=2n+1$)}
	\end{equation}
	Moreover, for $N = 2n$, the irreducible space is unique, while for $N = 2n +
	1$, we have two inequivalent irreducible representations with the same
	dimension $d$.
\end{thm}

\paragraph{Representation Theory on Real Clifford Algebra}
The two papers \cite{Okubo1991a}\cite{Okubo1991} are great in this respect.
Their result used in this document could be summarized as:
\begin{thm}
    Let $N=p+q$.
	Consider $\mathrm{C\ell}_{p,q}(\R)$. For $N=2n$, we have:
	\begin{align}
	d= \begin{cases}
	2^n_1, & \text{for } p-q=0\text{ or }2\text{, mod }8 \\
	2^{n+1}_1, & \text{for } p-q=4\text{ or }6\text{, mod }8 \\
	\end{cases}
	\end{align}
	For $N=2n+1$, we have:
	\begin{align}
	d= \begin{cases}
	2^n_2, & \text{for } p-q=1\text{, mod }8 \\
	2^{n+1}_1, & \text{for } p-q=3\text{ or }7\text{, mod }8 \\
	2^{n+1}_2, & \text{for } p-q=5\text{, mod }8 \\
	\end{cases}
	\end{align}
	Where we have used a subscript $1$ to denote that there is only
	one unique irreducible representation, and a subscript $2$ to
	denote that there are two inequivalent irreducible representations.
\end{thm}
Below collects a list of dimensions that I have used in the discussion of
Table~\ref{tab:mat-dim-allClass-1d}:
\begin{equation}
    \label{eq:Cl-dimenList}
    \begin{aligned}
    &
    \mathrm{C\ell}_{1,2}(\R)\sim 2^2_1,\, \mathrm{C\ell}_{1,3}(\R)\sim 4^2_1,\,
    \mathrm{C\ell}_{1,4}(\R)\sim 4^2_2,\, \mathrm{C\ell}_{1,5}(\R)\sim 8^2_1,\,
    \\&
    \mathrm{C\ell}_{2,2}(\R)\sim 2^2_1,\, \mathrm{C\ell}_{2,3}(\R)\sim 4^2_1,\,
    \mathrm{C\ell}_{3,2}(\R)\sim 2^2_2,\, \mathrm{C\ell}_{2,4}(\R)\sim 8^2_1,\,
    \\&
    \mathrm{C\ell}_{3,3}(\R)\sim 4^2_1,\, \mathrm{C\ell}_{4,2}(\R)\sim 4^2_1.
    \end{aligned}
\end{equation}