### 2017-05-16 Notes of various papers/Chiu Minimal Dirac Hamiltonian Method/chiu.SPEMT.tex

% The entire content of this work (including the source code
% for TeX files and the generated PDF documents) by 
% Hongxiang Chen (nicknamed we.taper, or just Taper) is
% licensed under a 
% Creative Commons Attribution-NonCommercial-ShareAlike 4.0 
% International License (Link to the complete license text:
% http://creativecommons.org/licenses/by-nc-sa/4.0/).
\documentclass{article}

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physicist-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}
\usepackage{fontspec}
\setmainfont{Times New Roman}

\makenomenclature % For an index of symbols.

\title{Notes about the Minimal Dimensional Method to Classify
Topological Phases (preliminary)}
\date{\today}
\author{Taper}

\begin{document}

\maketitle
\abstract{
    \todo{Haha, leave this until later.}
}
\tableofcontents

\input{chapters/introduction.tex}
\input{chapters/chap-2.tex}
\input{chapters/chap-3.tex}
\input{chapters/chap-4.tex}

\begin{appendix}
\input{chapters/chap-appA.tex}
\input{chapters/chap-appB.tex}
\end{appendix}

\bibliography{../../library}{}
\bibliographystyle{alphaurl}
% \begin{thebibliography}{1}
% 	\bibitem{book} 
% \end{thebibliography}
\printnomenclature

\end{document}