### 2018-03-10 Notes of various papers/ud-3-ortho-peres/ud-3-ortho.tex

\documentclass{article}

\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}
\usepackage[sharp]{easylist} % use # symbol to denote a list

\makenomenclature

\title{UD Between 3 Non-ortho}
\date{\today}
\author{Taper}

\begin{document}

\maketitle

\begin{abstract}
This is a note for \href{https://arxiv.org/abs/quant-ph/9804031}{Optimal distinction between non-orthogonal quantum states}.
\end{abstract}

\section{Background}

\begin{easylist}[enumerate]
# Why unambiguous?
## It may happen that we have unlimited supply of data
## But we cannot afford a mistake
### Example: cryptography (?)
### why cryptography? $\to$ {\color{red} don't know!}
# Nonortho for $2$ states: solved by [4-6]
## Chefles investigated $N$ linearly independent states $\to$ partial result
# Here: $3$ states
## uses POVM
## a failure could also give contribution using posterior probabilities
## could be extended to $N$ states $\to$ become tedious
\end{easylist}

\section{Setup}

\begin{easylist}[enumerate]
# $\dim{H} =3$
# Input: $u_1, u_2, u_3$, with \textit{a priori} probability $p_1,p_2,p_3$.
# Aim: maximize expected gain of signal
# Start with 
## something approximately $ v_1=u_2\times u_3$.
## it looks like orthogonal complement
## We will add restriction to $v_i$.
#Finally
## $A_j = k_j \ketbra{v_j}$, where $k_j$ is something we want to maximize.
## $P_j = \braket{u_j, A_j u_j}$, and $P_0 = \sum_{j=1}^3 p_j \braket{u_j, A_0 u_j}$.
## Obviously, the $k_j$ must make $A_0 =\id - \sum_{i=1}^3 A_j$ a valid positive operator
## The restriction is discussed in p.4 to 5, and especially eq (11)
## Most useful condition: $\det(A_0)\geq 0$, finally turn into: eq. (12)
# Result
## Let: $G=\sum_j C_j P_j$ , where $C_j$ is the value of signal for $u_j$.
## Maximal $G$ could be obtained by adjusting $k_j$, and the process is discussed in p.6 to p.7
## Example:
### When $p_j=\frac{1}{3}$, and all $C_j=1$, the optimal result is when $k_1=$ .... And $P_0=0.8386$.
### When $p_j$ = as above, but $C_1=0.8, C_2 = 1.2$ and $C_3=1$, the optimal result is in a different sets of $k_i$, and $P_0 = 0.8626$.
## {\color{red} It feels odd to talk about different values for $C_i$.}

# Why Inconclusive matters?
## The initial entropy is $H_\text{init} = -\sum_{j=1}^3 p_j \ln p_j$ 
## In general, the optimal $A_0$ is a matrix of rank $2$, written as: $A_0 = \lambda_m\ketbra{m} + \lambda_n \ketbra{n}$.
## The probability of $m$'s result is: $P_{mj} = p_j \lambda_m |\braket{m,u_j}|^2$
## Using Bayes' theorem, we could have, the posterior probability for input $u_j$ upon observing output$m$ is
### $Q_{jk} = P_{mj}/\sum_{i=1}^3 P_{mi}$ 
## The entropy of this $H_m = -\sum_{j=1}^3 Q_{jm} \ln Q_{jm}$ is usually $H_m < H_\text{init}$.
## Hence at least some information has been gained
## {\color{red} Is this a meaningful gain? How to make use of this?}
#  $N> 3$ signals
## only sketched, actually many are conjectures.

\end{easylist}
\end{document}